\section{Conclusion and Future Work}
\label{sec:conclusion}

We set out to replace vendor-benchmark claims with a neutral, reproducible
comparison of relational database access layers in Express.js. Our harness serves
an identical application through a representative taxonomy (native drivers, a query
builder, and six ORMs) across PostgreSQL and MySQL, over five access patterns,
and reports throughput and tail latency together. This addresses the four gaps that
characterize prior art: the absence of MySQL from access-layer comparisons, the
lack of a neutral, broad-coverage study, the separation of throughput from tail
latency, and the omission of the Express HTTP layer.

The empirical answer is that the cost of abstraction is concentrated, not uniform.
It is modest for point reads and for aggregation issued through a raw-SQL path,
but it grows sharply (up to $6.3\times$ between the native driver and the heaviest
data-mapper ORM) for the deep/nested fetch, where the layer must assemble a nested
object graph on the application side. A same-plan control decomposes that spread:
the layers' raw execution paths lie within about $1.4\times$ of one another, so most
of the deep-fetch cost is each layer's default eager-loading strategy and hydration
rather than its machinery. The coarse ordering of layers is stable
across the two engines, and a competitively fast modern ORM (Prisma) shows that
``ORM'' need not imply large overhead. Writes are the exception: under the relaxed-durability configuration used here,
MySQL's engine floor makes single-row insert throughput nearly layer-independent,
whereas PostgreSQL exposes a $4.7\times$ layer spread, a caution against generalizing
single-engine write results. Beyond these findings, development
surfaced a reusable checklist of benchmarking pitfalls (aggregation fan-out,
\texttt{OFFSET}-versus-keyset pagination, write-workload contamination, and
query-formulation confounds) that silently distort naive access-layer benchmarks;
one was caught by our correctness cross-check, the rest by performance analysis.

Beyond the single 50-connection operating point, we swept concurrency from 1 to 256
connections and confirmed the ordering is load-robust, sampled per-layer CPU and
memory footprint, and tested adjacent layers with Cliff's~$\delta$ and the
Mann--Whitney~$U$ test. Further extensions would still strengthen the evidence: a
fixed-rate, two-machine cross-run would bound coordinated-omission and
observer-interference effects; additional engines and dataset sizes would widen
external validity; and a dedicated study of the N+1 penalty (naive versus eager
loading per ORM) and of cold-start latency would deepen the practical guidance. The
complete replication package (harness, seed, adapters, and analysis) is released
under open licenses so that these extensions, and independent re-runs on current
library versions, are straightforward, addressing the repeatability gap widely
documented in computer-systems research~\cite{collberg2016repeatability}.
