\section{Discussion}
\label{sec:discussion}

\paragraph{Implications for practice.}
Read together, RQ1 and RQ3 point to the same conclusion: abstraction cost is not
a fixed, uniform overhead but is specific to the \emph{access pattern},
and it concentrates in the pattern that assembles a nested object graph. Two
estimands separate throughout this discussion: the \emph{default-configuration}
(idiomatic) effect that a practitioner gets from each layer's documented defaults, and
the \emph{controlled same-SQL} effect that remains once every layer executes
identical SQL. Under the default-configuration estimand, the deep/nested fetch
requires the access layer to hydrate rows into objects and resolve associations (the
object-relational impedance mismatch~\cite{ireland2009impedance} made concrete), not
merely to issue SQL, and this is where the data-mapper ORMs are most
expensive: Sequelize, TypeORM, Objection, and especially MikroORM (571~req/s) trail
the native \texttt{pg} driver (3{,}741~req/s) by up to $6.55\times$ on PostgreSQL, the
widest spread in the study (Table~\ref{tab:deep_fetch}), with a comparable
$4.50\times$ spread on MySQL. The tail shows the same pattern: MikroORM's deep-fetch
p99 is 106~ms against 20~ms for \texttt{pg} on PostgreSQL. A team whose hot path
assembles nested object graphs therefore pays the largest abstraction cost
on its highest-value traffic, and a fan-out sweep of the same endpoint over graphs of
0 to 500 comments shows the spread growing monotonically with graph size, from
$5.9\times$ at zero comments to $16.1\times$ at 500, with the leader flipping from
Prisma on small graphs to \texttt{pg} from 50 comments on, as hydration cost scales
with the number of materialized rows; the primary workload's roughly ten-comment
graph sits near the low end of that range. Point reads and aggregation are the opposite case:
dominated by connection and parsing cost that every layer shares, point reads spread
only $3.06\times$/$2.36\times$; aggregation spreads just $1.86\times$/$1.42\times$
once every adapter, native drivers included, runs the same raw
correlated-subquery plan, though Prisma's raw-SQL path still beats the slowest
layer, Objection, by $2.09\times$ (PostgreSQL) to $2.49\times$ (MySQL). On the
controlled same-SQL estimand, the idiomatic deep-fetch spread collapses to
$1.76\times$ on PostgreSQL and $2.21\times$ on MySQL
(Section~\ref{sec:results}, Table~\ref{tab:sameplan}), so the wide
default-configuration deficits lie upstream of raw SQL execution: in each layer's
query-generation and result-construction path (its default eager-loading strategy,
query count, and object hydration), which the control changes as a bundle rather
than isolating one component. MikroORM illustrates this most starkly: its default
path delivers only 0.20 of its own same-SQL throughput on PostgreSQL and 0.25 on
MySQL, a $5.0\times$/$4.1\times$ gap between its idiomatic and raw-SQL paths, while
Prisma's raw path is the fastest same-SQL cell on both engines. The same control exposes a new low end: Sequelize's raw-SQL
facility (\texttt{sequelize.query}) is slowest on both engines,
evidence of dispatch and marshalling overhead in its raw API rather than in query
strategy.
Round-trip count alone does not set throughput: Prisma and Objection both issue four
deep-fetch queries (statement counts tabulated in the online supplement), yet the two sit at opposite ends
of the default-configuration ranking. Knex and Drizzle sit in a consistent middle
ground on every pattern (never fastest, never among the slowest data-mapper ORMs),
which is what a query builder and a non-data-mapper, lightweight ORM should deliver.

\paragraph{Same SQL, different wire protocols.}
Server-side statement capture confirms that the same-SQL control's collapse is not
a protocol artifact: every layer emits byte-identical control statements on both
engines, and the wire protocols differ per layer as enumerated in
Section~\ref{sec:methodology}. Protocol choice alone does not explain the same-SQL
spread: Prisma leads it on MySQL while using the binary protocol, and
\texttt{mysql2-tuned}'s binary path sits within noise of plain \texttt{mysql2}'s
text path on most patterns.

\paragraph{The CPU cost behind parity.}
Prisma shows that an ORM need not be slow, consistent with evidence that ORM cost is
highly sensitive to implementation and tuning~\cite{vanzyl2009optimisations}. On the
deep fetch its compiled query engine is statistically equivalent to the native
driver: the paired difference is not significant (\texttt{pg} faster by a
geometric-mean ratio of $1.02$, paired permutation $p=0.058$), and a paired
two-one-sided-tests equivalence procedure against an a-priori $\pm5\%$ margin
accepts equivalence ($90\%$ CI $[1.004, 1.037]$; 3{,}685 vs.\ 3{,}741~req/s). A tuned, prepared-statement
baseline absent from earlier campaigns, \texttt{pg-tuned}, reaches 3{,}813~req/s
here, modestly ahead of both, so a tuned native ceiling is now measured
rather than assumed. Prisma also leads outright on the keyset range scan and the
raw-SQL aggregation on both engines (3{,}963/3{,}936 and 7{,}893/7{,}383~req/s), and
on the point read and deep fetch on MySQL (5{,}993 vs.\ \texttt{mysql2}'s 4{,}337;
3{,}701 vs.\ \texttt{mysql2-tuned}'s 2{,}436). That parity or leadership is not free:
at the \emph{same} ten-connection pool, Prisma's process draws roughly five times the
application-side CPU of every other layer on the deep fetch ($498\%$;
Table~\ref{tab:resources}, Figure~\ref{fig:cpu_tradeoff}), trading CPU for
latency~\cite{bonvoisin2024orm}, a cost a throughput-only benchmark would miss. How
large that cost is depends on which CPU one counts. By \emph{application}-tier CPU
the efficiency ranking inverts sharply (3{,}741 vs.\ 740~req per application-CPU-second
for \texttt{pg} and Prisma), but the native driver pushes about 3.6 database cores
of join work that this figure ignores; counting \emph{combined} application and
database CPU per request narrows \texttt{pg}'s advantage to about $1.4\times$
(813 vs.\ 596~req per combined-CPU-second; Supplement Table~S6). The honest reading
is therefore tier-specific: Prisma spends several times the application CPU of the
other layers, which matters directly when the application container's CPU is the
budgeted resource (a shared host, a per-core billing model), yet end-to-end it is
only modestly less compute-efficient than the native driver. The equal-CPU control
(Section~\ref{sec:results}) confirms the application-tier trade operationally: at an
equal one-core application budget Prisma is $4.1\times$ slower than \texttt{pg} and
needs about four cores to approach parity, while MikroORM barely moves with more
cores, confirming its deficit is a hydration cost rather than a parallelism-starved
one.

\paragraph{Queueing under load and the open-loop check.}
The pool sampler turns a previously inferential design property into a direct
measurement: on the deep fetch, the native PostgreSQL driver's ten-connection pool
sits fully utilized (pool\_used\_avg~$\approx 10$) with a mean of about 32
requests waiting in the acquisition queue (peaks of 39--40) under the 50
concurrent connections used throughout, confirming queueing rather than an idle pool. A
coordinated-omission-corrected, constant-arrival cross-run of the deep fetch on
PostgreSQL (online supplement) shows the closed-loop ranking transfers under
that stricter model: below each layer's capacity the corrected tail is flat and
small (\texttt{pg} and Prisma hold p99~$\leq$~2--4~ms up to 2{,}000~req/s, both
sustaining the full offered rate), but once the offered rate crosses a layer's
capacity the heavy data-mapper ORMs collapse rather than degrade gracefully,
Sequelize reaching only 772 of an offered 2{,}000~req/s with a third of the
issued requests timing out, and MikroORM collapsing already at 1{,}000~req/s (488
achieved), so a closed-loop benchmark alone underestimates how abruptly, not just
how much, a heavyweight ORM's tail degrades beyond capacity.

\paragraph{Writes: an engine-bound cost under default durability.}
RQ2 has a clear answer on inserts, and the headline figures now come from each
engine's \emph{default} durability configuration (Section~\ref{sec:methodology})
rather than the relaxed one used in earlier campaigns. Even with fsync on, PostgreSQL's insert
throughput remains layer-dependent: \texttt{pg-tuned} and \texttt{pg} lead at
6{,}280 and 6{,}064~req/s, Prisma and Knex stay competitive (5{,}465 and 4{,}693),
and MikroORM trails at 1{,}571, a $3.86\times$ spread (Table~\ref{tab:write}). At the
50-connection concurrency used throughout, group commit absorbs most of the
per-commit flush cost: a secondary, symmetrically relaxed run ($n=10$, both engines,
including MySQL's binlog sync off) shows relaxing durability buys native
PostgreSQL at most $14\%$ and at most a few percent on most layer-bound ORMs (median
$+8\%$ across the nine PostgreSQL cells; per-layer table in the online supplement), so the
layer-dependence is a property of the write path itself, not of fsync. MySQL
inserts, by contrast, remain low and flat across all nine layers (663--809~req/s, a
$1.22\times$ spread), the same ``engine floor'' pattern the relaxed-durability run
already showed; what default durability newly exposes is the mechanism in the tail:
MySQL's p99 sits at 135--155~ms across every layer, consistent with the per-commit
double flush (redo log, then binlog) InnoDB performs on every insert regardless of
how the request reaches it~\cite{harizopoulos2008oltp}, against 11--37~ms on
PostgreSQL. Relaxing MySQL's durability recovers only $19$--$24\%$, and the absolute
gap to PostgreSQL remains roughly $7\times$ even then (966 vs.\ 6{,}895~req/s
under the relaxed regime on both engines). The
engine contrast is not a durability artifact: under vendor defaults it is,
if anything, slightly larger than under the secondary, mechanism-isolating relaxed
run. This finding also warrants caution: a write-throughput comparison run on a
single engine invites the wrong generalization in either direction (``the access
layer is irrelevant to writes'' from MySQL alone, or ``the access layer matters a
great deal for writes'' from PostgreSQL alone), and an engine-only benchmark such
as~\cite{salunke2024pgmysql} cannot expose this interaction because it omits the
access layer entirely. Only the dual-engine design used here makes the
engine-write-path explanation visible instead of confounded with layer choice; that
the underlying database technology modulates the cost of a given mapping strategy is
consistent with earlier quantitative ORM work~\cite{lorenz2017orm}, and the ranking
itself is engine-specific: unlike every read pattern, the write ranking does not
transfer across engines (Spearman $\rho = 0.43$ versus $\rho \geq 0.89$ for reads).


\paragraph{Methodological lessons: a checklist for access-layer benchmarking.}
Reaching the figures above meant confronting four confounds that would otherwise
have silently corrupted the comparison. Their provenance differs. One, the
aggregation fan-out, is a \emph{correctness}
bug that our automated cross-check caught (the same check, which compares each
adapter's output against a native-driver reference on four of the five endpoints,
also caught a driver result-shape mismatch). The other three return \emph{correct}
results and were therefore invisible to any equality check; they were found only by
performance analysis. We regard the resulting checklist as reusable:
\begin{itemize}[leftmargin=*,noitemsep,topsep=2pt]
  \item \textbf{Aggregation fan-out.} A naive join between \texttt{posts} and
        \texttt{comments} evaluated before the \texttt{SUM} multiplies each
        post's view count by its number of matching comment rows, inflating
        the aggregate, a correctness bug invisible unless every layer's
        output is checked against a reference, not a performance one.
  \item \textbf{OFFSET vs.\ keyset pagination.} An early, OFFSET-based design
        for the range-scan endpoint collapsed on MySQL at
        realistic offsets, which would have been misread as an access-layer
        weakness rather than what it is: a pagination-strategy cost that
        InnoDB pays more dearly for than PostgreSQL. Rewriting every adapter
        around keyset pagination (\texttt{WHERE id < cursor ORDER BY id DESC
        LIMIT n}) removed the confound.
  \item \textbf{Write-workload contamination.} The insert endpoint grows
        \texttt{posts} on every call; because engines and layers commit at
        different rates, later cells in a run would scan a differently sized
        table depending on run order rather than on layer identity, unless
        benchmark-inserted rows are deleted before every cell.
  \item \textbf{Query-formulation confounds.} A first attempt at the
        aggregation endpoint pre-aggregated \emph{all} of \texttt{comments}
        before filtering to one author, so every request re-scanned the whole
        table regardless of which author was requested, a query-plan
        artifact masquerading as an access-layer effect~\cite{leis2015optimizers},
        fixed by rewriting
        the query as a correlated subquery touching only the target author's
        rows.
\end{itemize}
None of these four is specific to our schema; each is a generic failure mode
of the genre, extending the database-benchmarking pitfalls cataloged
by Fruth et al.~\cite{fruth2022tail} and Raasveldt et al.~\cite{raasveldt2018fair}
to the access-layer-comparison sub-genre specifically. We offer the checklist as a
practitioner aid for anyone comparing drivers, query builders, or ORMs: each confound
is easy to introduce by accident and hard to detect (the correctness ones without
an automated cross-check, the performance ones without profiling every cell), and at
least one such confound plausibly explains part of the gap between some
vendor-reported and independently verified figures in this space.

\paragraph{Practical guidance.}
The practical synthesis follows directly from RQ1--RQ3, read on the
default-configuration estimand that a practitioner faces. Where a service's
hot path is dominated by deep/nested fetches, the access layer is consequential and
a thin layer (the native driver, a query builder, or a query-engine-based ORM such
as Prisma) is the safer default \emph{for throughput and tail latency}, since a
heavyweight data-mapper ORM concentrates its largest penalty on that traffic, a
penalty that grows, not shrinks, as the fetched graph grows larger; weigh a
query-engine ORM's parity or lead here against its application-tier CPU cost
wherever CPU, not throughput, is the constrained resource. Where the
workload mix is read-simple (point reads, keyset scans) or write-bound, especially
on MySQL, whose engine floor for single-row inserts dominates layer differences
under default durability, the productivity, type safety, and
migration tooling of a full ORM cost little in relative throughput and are
reasonable grounds for the decision on their own. A pragmatic middle path, already
forced onto every layer by our aggregation endpoint, is to keep the ORM for the bulk
of the application and fall back to raw SQL or the native driver only for
endpoints that assemble large object graphs; the same-SQL control shows
this fallback recovers most of the deep fetch's cost (to at most $2.21\times$), and
orthogonally, second-level caching can absorb part of an ORM's read cost where the
workload tolerates staleness~\cite{chen2016cacheoptimizer}.

These are performance findings only: the access-layer decision also turns on
developer productivity, type safety, correctness, maintainability, and security,
none of which this study measures, and which may outweigh the throughput
differences reported here. Even as performance guidance, the recommendations rest on
one schema, one hardware configuration, a single-host local-database deployment
measured under each engine's default durability for writes, and the pinned library
versions; the same-SQL comparisons above decompose that cost, not a license to
replace idiomatic defaults with raw SQL in practice. A
production deployment adds network latency, TLS, a reverse proxy, a separate
database host, connection poolers, read replicas, and larger or more varied
payloads, any of which can shift the balance; Section~\ref{sec:threats} sets out how
far these results can be expected to generalize.
