\section{Introduction}
\label{sec:introduction}

Express.js remains one of the most widely used server frameworks for Node.js, and
almost every non-trivial Express service talks to a relational database. The way it
does so, however, is far from uniform. Developers choose among a spectrum of
\emph{access layers}: thin native drivers such as \texttt{pg} and \texttt{mysql2},
query builders such as Knex, and full object-relational mappers (ORMs) such as
Prisma, Sequelize, TypeORM, Objection, MikroORM, and the lighter-weight Drizzle.
Each layer trades raw performance for ergonomics, type safety, and protection from
foot-guns such as the N+1 query problem.

Practitioners typically justify this choice with \emph{vendor benchmarks}. As our
survey of prior art shows (Section~\ref{sec:related}), these are fragmented,
almost exclusively target PostgreSQL, and---critically---report either throughput
\emph{or} latency, seldom both, and rarely the tail (p95/p99) that dominates
user-perceived performance at scale~\cite{dean2013tail,fritz2021tail}. The
peer-reviewed literature is thinner still and covers at most two or three
libraries; the one cross-language academic study of ORM overhead deliberately
excludes JavaScript altogether~\cite{colley2018orm}.

This paper addresses that gap with a neutral, reproducible benchmark. We serve an
\emph{identical} Express application through nine access layers, over five
canonical access patterns, against \emph{both} PostgreSQL and MySQL, and we report
throughput and tail latency together.

\paragraph{Research questions.}
\begin{enumerate}[label=\textbf{RQ\arabic*},leftmargin=*]
  \item How much throughput and tail-latency overhead does each access layer add
        relative to the native-driver baseline, per access pattern?
  \item Does the ranking of access layers hold across PostgreSQL and MySQL, or is
        it engine-dependent?
  \item For which access patterns (point read, range scan, deep/nested fetch,
        aggregation, insert) is the layer choice most consequential?
\end{enumerate}

\paragraph{Contributions.}
(i)~A full-taxonomy, dual-engine, Express-specific benchmark harness released as an
open replication package; (ii)~a joint throughput/tail-latency comparison that
existing benchmarks omit; (iii)~the first inclusion of MySQL alongside PostgreSQL in
this comparison genre.
