\section{Study Design}
\label{sec:methodology}

Our goal, in Goal--Question--Metric terms, is to \emph{analyze} the relational
database access layer of an Express.js service \emph{for the purpose of}
quantifying its performance cost \emph{with respect to} throughput and tail
latency \emph{from the point of view of} an application developer choosing a layer
\emph{in the context of} PostgreSQL and MySQL back ends. The design crosses the
seven portable access layers with both engines and all five access patterns, plus
the two engine-specific native-driver baselines; the complete rationale, including
the pitfalls each control guards against, is distributed with the replication
package (\texttt{METHODOLOGY.md}).

\subsection{Factors and treatments}
We vary three factors. The \emph{access layer} has nine levels spanning the
taxonomy: the native drivers \texttt{pg} and \texttt{mysql2}, the query builder
Knex, and the ORMs Drizzle (a lightweight, query-builder-style ORM), Prisma,
Sequelize, TypeORM, Objection, and MikroORM. The \emph{engine} has two levels,
PostgreSQL~18.4 and MySQL~9.7.1. The \emph{access pattern} has five levels,
realized as HTTP endpoints (Table~\ref{tab:patterns}). The native drivers are
engine-specific (\texttt{pg} runs only on PostgreSQL, \texttt{mysql2} only on
MySQL); the other seven layers run on both engines, giving
$2 + 7\times 2 = 16$ layer\,$\times$\,engine cells and $16\times 5 = 80$
measured configurations.

\begin{table}[htbp]
  \centering
  \caption{The five access patterns and what each stresses.}
  \label{tab:patterns}
  \begin{tabular}{l l l}
    \toprule
    Endpoint & Pattern & Stresses \\
    \midrule
    \texttt{GET /posts/:id}          & point read (PK)       & per-query overhead \\
    \texttt{GET /posts?before=}      & keyset range scan     & index seek + hydration \\
    \texttt{GET /posts/:id/thread}   & deep/nested fetch     & join strategy, N+1 avoidance \\
    \texttt{GET /authors/:id/summary}& aggregation           & grouped counts / raw SQL \\
    \texttt{POST /posts}             & insert                & write path \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Objects: schema, workload, and data}
The workload is a minimal blog domain---\texttt{authors} $1\!-\!*$ \texttt{posts}
$1\!-\!*$ \texttt{comments} $*\!-\!1$ \texttt{authors}---which exercises a
one-to-many relationship and a two-hop join for the deep fetch while remaining
small enough to reason about. The five endpoints map onto the canonical access
patterns a CRUD service issues: a primary-key point read; a newest-first
\emph{keyset} page (\texttt{WHERE id < cursor ORDER BY id DESC LIMIT}~$n$, using
the primary-key index rather than a large \texttt{OFFSET}); a deep/nested fetch
that returns a post together with its author and every comment, each with its own
comment-author; a per-author aggregation (post count, comment count, total views);
and a single-row insert. The database is loaded from a deterministic seed
(2{,}000 authors, 100{,}000 posts, 1{,}000{,}000 comments) by the native driver,
independently of any layer under test, so that every engine receives byte-identical
data. The working set fits in memory, so measurements reflect access-layer cost
rather than storage latency.

\subsection{The adapter contract and N+1 control}
Every access layer implements the same factory interface---five query methods plus
a teardown---so the Express server and the load runner are layer-agnostic. Each
adapter returns an \emph{identical result shape}, and each uses that layer's
\emph{recommended idiomatic} mechanism to avoid the N+1 query problem on the deep
fetch: a hand-written join for the native drivers, the Knex query builder, and the
query-builder-style Drizzle, and relation eager loading (\texttt{include},
\texttt{withGraphFetched}, \texttt{populate}, or \texttt{relations}) for the
data-mapper ORMs. To guarantee that no
adapter gains an unfair advantage by silently returning less---or by issuing N+1
queries---an automated cross-check asserts that all nine adapters produce the same
normalized object for the same inputs before any timing is collected. This check
caught two genuine defects during development (a fan-out aggregation bug and a
driver result-shape mismatch; see Section~\ref{sec:discussion}).

\subsection{Controls}
Four controls isolate the access layer as the only variable. \emph{(i)~Pool size}
is fixed at ten connections for every adapter, so the comparison measures
access-layer overhead rather than pool tuning, itself a known web-application
performance factor~\cite{gupta2017pooling}; for Prisma, whose default pool is
otherwise derived from the core count, this is pinned explicitly through the
\texttt{connection\_limit} URL parameter. \emph{(ii)~Identical query
semantics}: each endpoint issues the same logical query across layers; the
aggregation, in particular, uses a single correlated-subquery formulation
everywhere (the documented raw-SQL escape hatch), so that endpoint measures layer
overhead on an identical plan rather than differences in generated SQL.
\emph{(iii)~Write isolation}: because the insert endpoint mutates the dataset, the
runner deletes benchmark-inserted rows before every cell, so each cell starts from
the identical seeded table. \emph{(iv)~Observer separation}: the HTTP load
generator runs in a process separate from the server.

\subsection{Measurement procedure}
Load was generated with \texttt{autocannon}~\cite{autocannon}, an HTTP/1.1 load
tester that issues requests continuously at a fixed concurrency and records a
latency histogram; it is a closed-loop (constant-concurrency) generator. For each
of the 80 configurations we opened 50 concurrent connections, ran a two-second
warm-up pass whose measurements were discarded (absorbing JIT compilation,
connection-pool fill, and query-plan caching), then took ten measured four-second
runs and reported the median. Each run yields throughput (requests per second) and
latency percentiles p50, p90 and p99. During the measured runs we also sampled the
\emph{server} process's CPU utilization and peak resident memory from
\texttt{/proc}; because the load generator runs in a separate process, this isolates
the access layer's own footprint from the load tool and the database. Both engines
ran with the working set resident in memory and durability relaxed, so the results
reflect access-layer CPU and allocation cost rather than disk synchronization. The
hardware and software environment was captured automatically into the replication
package. To characterize behavior under load, we additionally swept the deep/nested
fetch---the most layer-sensitive pattern---across connection counts from 1 to 256
for every layer and engine.

\subsection{Analysis}
Because absolute throughput is hardware-specific, we analyze \emph{relative}
differences across layers within an engine. We quantify a pattern's sensitivity to
the access layer by its \emph{spread}: the ratio of the native driver's median
throughput to that of the slowest layer on that pattern (native-relative, so it is
comparable across patterns even where one ORM exceeds the native driver). We report
medians over the ten
runs and the coefficient of variation (CV) across runs as a stability signal,
following the practice of reporting central tendency and variability rather than a
single number~\cite{georges2007rigorous} (Section~\ref{sec:results}). To test whether
adjacently ranked layers are
distinguishable rather than an artifact of measurement noise, we compare each
adjacent pair on the deep/nested fetch with the Mann--Whitney $U$ test (two-sided)
and report Cliff's~$\delta$ as a nonparametric effect size, the latter being robust
to the small number of runs and to non-normal throughput distributions. We do not
apply significance testing to widely separated layers, where the effect is obvious;
the tests target the close pairs where a ranking could plausibly be noise.

\subsection{Experimental setup}
All measurements were taken on a single host (Intel Xeon Gold~6140, 16~cores,
126~GB RAM; Linux~6.17; Node.js~24.16.0) with the database engines running locally
under relaxed durability. Because access-layer performance is version-sensitive, we
pin every dependency (Table~\ref{tab:versions}); these are the versions the reported
numbers characterize, and re-running the harness on newer releases is a one-command
operation.

\begin{table}[htbp]
  \centering
  \caption{Pinned versions of the evaluated access layers and tooling.}
  \label{tab:versions}
  \begin{tabular}{l l l l}
    \toprule
    Layer / tool & Version & Layer / tool & Version \\
    \midrule
    \texttt{pg}         & 8.22.0  & \texttt{sequelize}      & 6.37.8 \\
    \texttt{mysql2}     & 3.22.5  & \texttt{typeorm}        & 0.3.30 \\
    \texttt{knex}       & 3.3.0   & \texttt{objection}      & 3.1.5  \\
    \texttt{drizzle-orm}& 0.36.4  & \texttt{@mikro-orm/core}& 6.6.15 \\
    \texttt{@prisma/client} & 5.22.0 & \texttt{express}     & 4.22.2 \\
    \texttt{autocannon} & 7.15.0  & PostgreSQL / MySQL      & 18.4 / 9.7.1 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Replication package}
The harness---the Express application, the adapter contract and all nine adapters,
the deterministic seed, the \texttt{autocannon} matrix runner, the correctness
cross-check, and the scripts that regenerate every table in
Section~\ref{sec:results}---is released so the full matrix can be re-run with a
single command against a containerized or local PostgreSQL and MySQL.
