\section{Related Work}
\label{sec:related}

We survey prior work along four \emph{coverage} axes --- taxonomy, engine, joint throughput/tail
reporting, and vendor involvement --- our case study spans (Table~\ref{tab:prior_art}); it leaves the
comparability of its numbers unestablished, the gap our protocol addresses. A structured
scoping search (not a systematic review) of the ACM Digital Library, IEEE Xplore, Scopus, and Google
Scholar (2006--July~2026) retained empirical comparisons of relational access layers or engines, with
search details archived in the replication package (\texttt{notes/related-work-search.md}); being
scoping rather than exhaustive, we scope every gap claim to the sources searched.

\subsection{Object-relational mapping and the access-layer spectrum}
Access layers bridge the object-relational impedance mismatch between the relational model and an
application's object graphs~\cite{ireland2009impedance}, spanning native drivers, query builders, and
ORMs following Data Mapper / Active Record patterns~\cite{fowler2002patterns}
(Section~\ref{sec:introduction}). Torres et al.~\cite{torres2017orm} survey two decades of ORM
patterns; their trade-off --- productivity against a hard-to-predict runtime cost --- is what we
quantify, holding the store choice (relational vs.\ non-relational~\cite{sadalage2012nosql}) fixed
while varying only the access layer.

\subsection{ORM performance and data-access anti-patterns}
A parallel line studies ORM performance in reverse, repairing the inefficient SQL frameworks emit. The
costliest is the N+1 query problem (Section~\ref{sec:introduction}); it and related redundant-query
anti-patterns pervade real-world web
applications~\cite{yang2018dbapps,yan2017inefficiencies,shao2020antipatterns}, their impact quantified
in ORM-backed systems~\cite{chen2016redundant,chen2014orm} and often \emph{removable} by query
synthesis, batching, or caching~\cite{cheung2013querysynthesis,cheung2014sloth,chen2016cacheoptimizer}
(with a JavaScript \texttt{async}/\texttt{await} analogue~\cite{turcotte2022asyncjs}). Closest to our
design, Lorenz et al.~\cite{lorenz2017orm} show the best mapping strategy depends on database technology,
van Zyl et al.~\cite{vanzyl2009optimisations} that it is tuning-sensitive, and a recent study adds an
energy dimension~\cite{bonvoisin2024orm}. Overhead is thus configurable, not fixed --- as
our per-pattern results confirm --- but this line never compares drivers, query builders, and ORMs
head-to-head under load, our aim.

Cross-implementation ORM equivalence checking itself has a direct precedent. CYNTHIA generates an
abstract query, translates it to equivalent queries for five ORMs (including Sequelize), executes
them over four relational DBMSs, and differentially compares the results to find correctness
bugs~\cite{sotiropoulos2021cynthia}. It therefore shows that semantically aligned ORM operations
can serve as a cross-implementation differential oracle. CYNTHIA is a testing method, however, not a
performance-benchmark admission protocol: it does not use equivalence to decide which timing cells
may enter a benchmark, define a premeasurement rule for which implementation represents each
library, or separate equal-demand capacity effects from matched-utilization latency. Our novelty
claim is restricted to that joint operationalization.

\subsection{Peer-reviewed access-layer comparisons}
Peer-reviewed head-to-head access-layer measurement is scarce. Colley et al.~\cite{colley2018orm} give
the earliest cross-language account --- eight ORMs across Java, C\#, Python, and PHP on a common schema
--- but \emph{deliberately exclude JavaScript}, leaving it without an academic baseline for nearly a
decade. Three recent studies partially address this: one in
\emph{Procedia Computer Science} compares Prisma against optimized raw SQL over eight query patterns on
PostgreSQL~\cite{prisma_rawsql_2025} (differing versions, hardware, and workload, so we neither build on
nor dispute its figures); a second compares Prisma, TypeORM, and Sequelize across four relation types,
also on PostgreSQL, by response time, memory, and CPU~\cite{orm_compare_jcct_2025}; the most recent times
internal query stages of the same three ORMs rather than client-observed
requests~\cite{zhadko2025orm}, independently observing the concurrency-dependent Prisma ranking flip we
also report. A further Express study optimizes one stack rather than comparing access
layers~\cite{imamsakti2025async}.

\subsection{Database-engine, SQL/NoSQL, and Node.js performance}
A second, disjoint line benchmarks the layers below and above the access layer without varying it.
Salunke and Ouda~\cite{salunke2024pgmysql} compare PostgreSQL and MySQL under standard OLTP
workloads (the only dual-engine precedent found), but benchmark the \emph{engines themselves}, so cannot
say how an access layer changes it; SQL/NoSQL comparisons~\cite{li2013sqlnosql} likewise measure the
store, not the layer. On the Node.js side, work establishes Node's web
viability~\cite{chaniotis2015nodejs}, tracks back-end drift under sustained load~\cite{kuffel2025nodejs},
and compares frameworks broadly~\cite{carmo2024backend}, but none treats the access layer as an
independent variable across the taxonomy studied here.

\subsection{Standard benchmarks and benchmarking methodology}
Standard benchmarks define how systems are measured: YCSB~\cite{cooper2010ycsb} for cloud-serving
stores and TPC-C/TPC-E for OLTP. These target the \emph{engine}, not the driver, query builder, or ORM
this study varies, but set the bar for controlled, repeatable load. A methodological literature
explains why single-metric reporting fails: Fruth et al.~\cite{fruth2022tail} and Raasveldt et
al.~\cite{raasveldt2018fair} catalog benchmarking pitfalls (coordinated omission, warm-up); Dean and
Barroso~\cite{dean2013tail} show the tail, not the mean, drives user-perceived latency under fan-out;
and Li et al.~\cite{li2014tales} trace its sources. Such measurement is often ad
hoc~\cite{jiang2015load,leitner2017exploratory} and hard to reproduce~\cite{collberg2016repeatability},
motivating a harness built for relevance, repeatability, and
fairness~\cite{huppler2009benchmark,sim2003benchmark}. We take a choice absent above: report throughput
\emph{and} p99 for the same run, not a single after-the-fact metric.

\subsection{Practitioner and vendor benchmarks}
A larger body comes from practitioners and vendors, each authored by a party with a stake in a compared
product. Drizzle publishes a Northwind micro-suite (nine targets) and a k6 load test reporting req/s,
average latency, P95, and CPU~\cite{drizzle_benchmarks}. Prisma's site
compares Prisma, Drizzle, and TypeORM by median query latency over 500 iterations, reporting neither
throughput nor any percentile~\cite{prisma_benchmarks}. IMDBench adds Sequelize and node-postgres,
reporting throughput and latency on three CRUD operations~\cite{imdbench}. Together these cover more
taxonomy than the peer-reviewed literature combined, yet share the defects of
Section~\ref{sec:introduction}: vendor authorship, PostgreSQL-only coverage, and a single metric family.

\subsection{Positioning}
In the sources searched, access-layer benchmarks report numbers whose comparability is assumed, not
established, and we found no study combining broad taxonomy, both engines, joint throughput/tail
reporting, and non-vendor authorship --- each covers at most two or three
(Table~\ref{tab:prior_art}), the strongest claim the scoping search supports, not one of priority. But
coverage is not the contribution: the protocol operationalizes established experimental principles for
this genre, not a new one. Kounev et al.'s systems-benchmarking
criteria~\cite{kounev2020benchmarking}, Hasselbring's software-engineering
standard~\cite{hasselbring2021benchmarking}, and Raasveldt and M\"uhleisen's fair-benchmarking
guidance~\cite{raasveldt2018fair} supply its principles at a deliberately general level. They leave an
access-layer study to operationalize \emph{what code represents a library}, how competing
implementations are checked for semantic equivalence, and how strategy-sensitive results are
contrasted. Our protocol supplies one concrete operationalization: a documentation-selected,
predeclared treatment rule; a same-SQL diagnostic; and a per-cell gate requiring equivalent responses
and verified write state \emph{before} timing. Correctness-before-timing has precedents (SPEC validation, TPC ACID, and SQLancer~\cite{rigger2020sqlancer}), while
cross-implementation equivalence is also prior art: CYNTHIA constructs equivalent ORM queries and
compares their results as a differential oracle~\cite{sotiropoulos2021cynthia}. We do not claim to
invent that check. Our narrower contribution uses it as an explicit prerequisite for admitting
benchmark timings and combines it with predeclared treatment selection and operating-point
separation. Together these constructs --- predeclared treatment-selection,
cross-implementation admission-gated equivalence, and a same-SQL diagnostic --- yield one
domain-specific property general frameworks leave to the practitioner: an admission-and-interpretation
discipline making \emph{competing
implementations of one task} mutually comparable. That domain specialization, realized in the open
harness (Section~\ref{sec:introduction}), is the contribution, not the invention of the underlying
principles.

\begin{table}[htbp]
  \centering
  \small
  \caption{Prior work on relational-database access-layer performance,
    positioned along four \emph{coverage} axes our case study spans; the paper's contribution
    is the comparability protocol of Section~\ref{sec:introduction}, not this coverage.
    ``Layers covered'' aggregates native drivers, query builders, and ORMs
    into one count; ``none'' marks studies that vary the engine or
    framework but not the access layer. \emph{Vendor involvement} marks maintainer-authored
    studies --- a conflict-of-interest property of the source, not a methodological guarantee that
    a non-vendor-authored benchmark is unbiased.}
  \label{tab:prior_art}
  \begin{adjustbox}{max width=\textwidth}
  \begin{tabular}{@{}llllc@{}}
    \toprule
    Study & Layers covered & Engines & Metrics & Vendor involvement \\
    \midrule
    Colley et al.~\cite{colley2018orm} & 8 ORMs, 4 languages (no JS) & not stated & query latency & None \\
    Procedia CS~\cite{prisma_rawsql_2025}$^\dagger$ & Prisma vs.\ raw SQL & PostgreSQL & latency, CPU & None \\
    JCCT~\cite{orm_compare_jcct_2025} & 3 ORMs & PostgreSQL & latency, memory, CPU & None \\
    JCSI~\cite{zhadko2025orm} & 3 ORMs & PostgreSQL & per-stage latency & None \\
    Salunke and Ouda~\cite{salunke2024pgmysql} & none (engine only) & PostgreSQL + MySQL & throughput, latency & None \\
    Drizzle~\cite{drizzle_benchmarks} & 9 (broad) & PostgreSQL & throughput, latency, P95 & Vendor \\
    Prisma~\cite{prisma_benchmarks} & 3 ORMs & PostgreSQL & median latency only & Vendor \\
    IMDBench~\cite{imdbench} & 4 ORMs + driver & PostgreSQL & throughput, latency & Vendor \\
    \midrule
    \textbf{This study} & \textbf{9 documentation-selected (+2 tuned)} & \textbf{PostgreSQL + MySQL} & \textbf{throughput + response-time p99} & \textbf{None} \\
    \bottomrule
  \end{tabular}
  \end{adjustbox}
  \vspace{4pt}
  {\raggedright\footnotesize $^\dagger$Cited for its methodology only; its absolute speedup figures are not comparable to ours given the differing versions, hardware, workload, and harness.\par}
\end{table}
