\section{Results}
\label{sec:results}

The tables below report throughput (requests per second, higher is better) and
tail latency (p99, milliseconds, lower is better) per access layer and engine,
one table per access pattern. All numbers come from the run described in
Section~\ref{sec:methodology}: 25 independent replicates per cell, each a
freshly booted server process, pool fixed at ten, inserts under each engine's
default durability configuration. Each native driver is reported both
idiomatically (\texttt{pg}, \texttt{mysql2}) and with reusable prepared
statements (\texttt{pg-tuned}, \texttt{mysql2-tuned}); the tuned rows are a
practitioner-oriented lower bound on what the wire protocol allows under this
workload, not a ceiling every idiomatic layer could reach without code
changes. Absolute throughput is hardware-specific, so the comparison of
interest is \emph{relative} across layers within an engine
(Section~\ref{sec:methodology}). Two estimands recur below: the
\emph{default-configuration (idiomatic) effect} (the primary estimand behind
RQ1--RQ3, realized in the five tables above), and the \emph{same-SQL effect},
what remains when every layer executes identical SQL with identical row
mapping (Table~\ref{tab:sameplan}); the two are reported separately below.

% One combined table per access pattern: throughput (req/s) and tail latency (p99)
% side by side, per access layer and engine. \tabledir is set by each main file
% (tables/ for the generic build in paper/, ../tables/ for the IST build in
% paper/ist/), so this shared section resolves the path without relying on TEXINPUTS.
% (p50/p90 are in the replication package's raw results.)
\input{\tabledir point_read}
\input{\tabledir range_scan}
\input{\tabledir deep_fetch}
\input{\tabledir aggregation}
\input{\tabledir write}

\subsection{RQ1: overhead relative to the native baseline}
The native \texttt{pg} driver, or its tuned counterpart, leads only on the
PostgreSQL point read and insert. On every other pattern-engine combination,
Prisma leads outright, including over the tuned native baseline: the
keyset range scan on both engines, aggregation on both engines, the point read on
MySQL, and the deep fetch on MySQL (Tables~\ref{tab:point_read},
\ref{tab:range_scan}, \ref{tab:deep_fetch}, \ref{tab:aggregation}). It does so at
roughly five times the application-side CPU of every other layer ($498\%$ on the
deep fetch; Supplement Table~S10), so the finding is a CPU-for-throughput
trade, not a free win. The remaining PostgreSQL deep fetch is close enough
between \texttt{pg} and Prisma to need a formal equivalence test (below). The
cost of abstraction remains strongly \emph{pattern}- and, now visibly,
\emph{engine}-dependent.

The tuned baselines sharpen what ``native'' means as a ceiling. On
PostgreSQL, \texttt{pg-tuned} tops both the point read (6{,}851 vs.\
\texttt{pg}'s 6{,}695~req/s) and the insert (6{,}280 vs.\ 6{,}064), and raises the
deep-fetch ceiling slightly further (3{,}813 vs.\ 3{,}741); reusable prepared
statements are worth a few percent on PostgreSQL. On MySQL, \texttt{mysql2-tuned}
is within noise of idiomatic \texttt{mysql2} on the point read, range scan,
aggregation, and insert (differences of at most $0.8\%$), except on the deep
fetch, where it reaches 2{,}436~req/s against \texttt{mysql2}'s 2{,}405: the
binary protocol and server-side prepared statements (wire-protocol evidence
below) pay off only once a request issues more than one round trip.

On the \textbf{point read}, PostgreSQL ranges from \texttt{pg-tuned}'s 6{,}851
down to MikroORM's 2{,}185~req/s; the native-relative spread (\texttt{pg}'s
idiomatic 6{,}695 divided by the slowest layer) is $3.06\times$ (95\% CI
[2.97--3.16]). MySQL's spread is smaller, $2.36\times$ (\texttt{mysql2} 4{,}337
down to MikroORM 1{,}839; CI [2.31--2.42]), but MySQL's fastest layer is not the
native driver: Prisma leads at 5{,}993~req/s, $38\%$ above \texttt{mysql2}.

On the \textbf{deep/nested fetch} (the pattern that assembles a post, its author,
and all comments with their authors), the native-relative spread is the widest
measured: $6.55\times$ on PostgreSQL (\texttt{pg} 3{,}741 down to MikroORM's
571~req/s; CI [6.46--6.65]) and $4.50\times$ on MySQL (\texttt{mysql2} 2{,}405
down to MikroORM's 535; CI [4.41--4.59]). Prisma sits close behind \texttt{pg} on
PostgreSQL (3{,}685) and ahead of \texttt{mysql2} on MySQL (3{,}701,
$54\%$ above native). Because every layer runs the identical
per-replicate request stream, the comparison is paired (Section~\ref{sec:methodology});
\texttt{pg} is faster in 16 of 25 replicates, by a geometric-mean ratio of
$1.02$ (paired 95\% CI [1.00--1.04]; Table~\ref{tab:significance}), a difference
the paired sign-flip permutation test does not resolve as significant
($p=0.058$; the Wilcoxon signed-rank test agrees, $p=0.08$). A paired
two-one-sided-tests (TOST) equivalence test at an a-priori $\pm5\%$ margin
confirms the two are equivalent (ratio $1.02$, 90\% CI [1.004--1.037], inside the
$\pm5\%$ band), and Prisma's p99 is the lower of the two (18~ms vs.\ 20~ms). We
therefore read the top of the PostgreSQL deep-fetch ranking as a practical tie.
Knex (2{,}491 / 2{,}050) and the lightweight ORM Drizzle (2{,}080 / 1{,}446)
remain competitive on both engines, whereas the data-mapper ORMs trail by up to
$6.55\times$ on PostgreSQL: TypeORM, Sequelize, Objection, and especially
MikroORM.

\textbf{Aggregation} overhead shrinks once every layer is forced onto the same
raw correlated-subquery plan, but Prisma leads it outright on both engines,
including over the tuned native baseline: 7{,}893~req/s on PostgreSQL (ahead of
\texttt{pg-tuned}'s 7{,}731 and \texttt{pg}'s 7{,}035) and 7{,}383 on MySQL,
$76\%$ above \texttt{mysql2}'s 4{,}195. Excluding Prisma, the native-relative
spread (\texttt{pg} or \texttt{mysql2} divided by the slowest layer, Objection on
both engines) is $1.86\times$ on PostgreSQL (CI [1.83--1.89]) and $1.42\times$ on
MySQL (CI [1.40--1.44]). Abstraction is thus inexpensive when a layer merely
forwards a query, and on this pattern and the keyset range scan, Prisma's query
engine turns that low cost into an outright lead; abstraction is
costly only when the layer must materialize a nested object graph on the
application side.

\subsection{Same-SQL control}
The idiomatic deep fetch lets each layer choose its own query strategy, so its
spread mixes raw SQL execution with everything a layer does around it. To bound
the raw-execution component, we re-measured the deep fetch under a
\emph{same-SQL control}: every layer executes identical two-statement SQL (and,
per engine, identical \texttt{EXPLAIN} plans) with identical row mapping through
its raw-SQL facility, as the aggregation endpoint already does
(Table~\ref{tab:sameplan}). Server-side capture confirms the statement text is
byte-identical across layers on both engines, while the wire protocols differ
per layer as enumerated in Section~\ref{sec:methodology}; the honest description
of this control is therefore \emph{same SQL, protocol disclosed}, not
\emph{same plan}. The contrast is composite: switching from a layer's idiomatic
relational-fetch path to the common raw-SQL-and-shared-mapping path changes
its query strategy, query count, prepared-statement behavior, the raw versus
ORM API, and result marshalling together, so it isolates none of these
individually; what it establishes is that the raw-execution component is
small and the idiomatic overhead lies upstream of it.

On the identical SQL, the deep-fetch spread collapses to a max/min ratio
(across the nine layers; the idiomatic spreads above are native-relative, whose
max/min counterparts are $6.7\times$ and $6.9\times$) of $1.76\times$ on PostgreSQL (top: Prisma's raw path at 4{,}342~req/s;
bottom: Sequelize's raw facility at 2{,}472) and $2.21\times$ on MySQL (Prisma
4{,}220; Sequelize 1{,}912). Prisma's raw path leads on both engines, consistent
with its pipelined query engine; the new low end is Sequelize's raw facility
(\texttt{sequelize.query}), which we attribute to dispatch and marshalling
overhead in the raw API itself, not SQL execution. MikroORM shows the largest
strategy gap: its idiomatic path (571~req/s on PostgreSQL, 535 on MySQL) reaches
only 0.20 and 0.25 of its own same-SQL throughput (2{,}868 and 2{,}172), a
$5.0\times$ and $4.1\times$ gap between its idiomatic and raw-SQL paths. On this
pattern, the access layer's idiomatic cost is dominated by what it \emph{chooses to
do} by default (its query strategy, round-trip count, and object hydration), not
by the speed of its raw SQL execution; the fixed-payload Express-plus-JSON
floor (Section~\ref{sec:methodology}) sits well above every same-SQL cell, so
the framework floor does not manufacture this collapse.

\input{\tabledir sameplan}

\subsection{RQ2: engine dependence of the ranking}
The relative ordering of access layers is largely, but not uniformly, stable
across engines. The Spearman rank correlation between the PostgreSQL and MySQL orderings of
the seven portable layers (reported descriptively, since these seven are the
evaluated systems rather than a sample from a population;
Section~\ref{sec:methodology}) is $\rho=0.89$ on the point read, $0.89$ on the
range scan, $0.96$ on the deep fetch, and $0.89$ on aggregation (Kendall's
$\tau$ between 0.81 and 0.90 on all four), so a layer's rank on one engine
tracks its rank on the other for every read pattern.
The \textbf{insert} ranking is the exception ($\rho=0.43$, $\tau=0.43$). Prisma
illustrates why: it is the fastest ORM on PostgreSQL inserts (5{,}465~req/s, ahead
of Objection's 3{,}926) but the second-slowest layer on MySQL inserts (714, ahead
of only MikroORM's 663), a rank flip a benchmark run on either engine alone would
not reveal. Beyond order, the \emph{magnitude} of a layer's cost also depends on the
engine: a blocked permutation test on the paired PostgreSQL-minus-MySQL
log-throughput difference (Section~\ref{sec:methodology}) finds the
layer$\times$engine interaction significant on all five patterns ($p \leq
5\times10^{-5}$, the permutation-resolution floor, in each case), so
single-engine results are not portable in degree even where they are
portable in order.

Prisma's engine-dependent lead is visible on the \textbf{keyset range
scan} (Table~\ref{tab:range_scan}): it leads on both engines (3{,}963 on
PostgreSQL, 3{,}936 on MySQL), ahead of \texttt{pg-tuned} (3{,}811), \texttt{pg}
(3{,}782), and \texttt{mysql2} (3{,}231). The native-relative spread is
$4.89\times$ on PostgreSQL (CI [4.80--4.99]) and $4.43\times$ on MySQL (CI
[4.35--4.54]), confirming that the earlier collapse observed under
\texttt{OFFSET} pagination was a workload artifact rather than an engine or
layer property (Section~\ref{sec:discussion}). Most of that spread traces to a
single layer: MikroORM's range-scan throughput is 773~req/s on PostgreSQL and
729 on MySQL (p99 of 74 and 79~ms), well below every other layer on either
engine, an anomaly we flag but do not fully isolate on this endpoint; it is
consistent with the per-request cost of materializing the returned page into
MikroORM's identity map, the same hydration overhead the same-SQL control
isolates on the deep fetch (ratio 0.20 on PostgreSQL, above).

\textbf{Inserts} are the pattern where the engine dominates most visibly, and
where the durability regime matters to that claim. All headline write numbers
here are measured under each engine's \emph{default} durability configuration
(Section~\ref{sec:methodology}). On MySQL, insert throughput is low and
tightly clustered across all nine layers (663--809~req/s, a $1.22\times$ spread,
CI [1.20--1.24]; Table~\ref{tab:write}) and its p99 sits at roughly 135--155~ms for every layer (a
per-commit redo-log and binlog flush), whereas PostgreSQL spreads $3.86\times$
(\texttt{pg-tuned} 6{,}280 down to MikroORM 1{,}571, CI [3.77--3.98]). A
default-versus-relaxed comparison ($n=25$ vs.\ $n=10$ medians;
Supplement Table~S3) shows this is not a durability artifact: at 50
concurrent connections, group commit amortizes most of the per-commit
flush cost, so relaxing durability buys at most $14\%$ on PostgreSQL (native;
roughly $0$--$4\%$ on the layer-bound ORMs, median $+8\%$ across all nine rows)
and $19$--$24\%$ on MySQL (median $+20\%$); if anything, the engine contrast is
slightly larger under default durability than under the relaxed configuration
used elsewhere. On MySQL, insert throughput
is therefore governed mainly by the engine rather than the access layer; on
PostgreSQL the access layer remains the larger factor. Which layer is
fastest, however, does not transfer between the two: the ranking, not just the
magnitude, is engine-dependent for writes in a way it is not for any read
pattern.

\subsection{RQ3: which patterns matter most}
Ranking the five patterns by native-relative spread on PostgreSQL gives deep
fetch ($6.55\times$) $>$ range scan ($4.89\times$) $>$ insert ($3.86\times$)
$>$ point read ($3.06\times$) $>$ aggregation ($1.86\times$). The same
deep~$>$~range~$>$~point~$>$~aggregation order holds on MySQL ($4.50\times$,
$4.43\times$, $2.36\times$, $1.42\times$); only the insert moves, from third
place on PostgreSQL to last on MySQL ($1.22\times$): the pattern where MySQL's
engine floor, not the access layer, sets the ceiling (RQ2 above). The
deep/nested fetch is the most consequential pattern on both engines; aggregation
is the least, once every layer runs the same raw-SQL plan (the range scan's
large nominal spread, as above, is driven disproportionately by MikroORM
alone).

The primary deep-fetch workload uses posts with roughly ten comments on
average, near the low end of a fan-out sweep across 0, 1, 10, 50, 100, and 500
comments (PostgreSQL, medians of three replicates) that shows the layer
spread widening with graph size: $5.9\times$, $5.8\times$, $7.1\times$,
$9.8\times$, $11.4\times$, and $16.1\times$ respectively; the leader flips
from Prisma at small graphs (up to 10 comments) to \texttt{pg} at 50 comments
and above, as hydration cost scales with the number of materialized rows, and
MySQL shows the same qualitative shape. The deep-fetch spreads reported above
are therefore conservative: a larger graph would widen the gap between the
lightest and heaviest layers further.

\subsection{Tail latency}
We report p99 for every pattern (the tail metric that governs perceived latency
under load); p50, p90, and p97.5 are in the replication package. Tail latency
tracks throughput: on the deep/nested fetch, PostgreSQL's p99 ladder runs
\texttt{pg} 20, Prisma 18, Knex 27, Drizzle 31, TypeORM 45, Objection 55,
Sequelize 60, and MikroORM 106~ms (Table~\ref{tab:deep_fetch}), the same order as
the throughput ranking, so a heavyweight ORM's nested-fetch penalty is visible
not only in mean capacity but in the tail under load, a quantity single-metric,
throughput-only vendor benchmarks omit.

These closed-loop p99 values come from a constant-concurrency generator. To
check that the tail ranking is not an artifact of that model, we drove the
deep/nested fetch on PostgreSQL under the coordinated-omission-corrected
open-loop harness described in Section~\ref{sec:methodology} (timed-out
requests, a 10~s hard cap, clipped into the distribution; full sweep in
Supplement Table~S6). Below saturation the corrected tails
are flat and small: \texttt{pg} and Prisma hold p99 at 2--4~ms up to
2{,}000~req/s, both sustaining the full offered rate, and Knex's p99 is still
only 33~ms at 2{,}000. Beyond each layer's capacity the tail collapses rather
than degrading gracefully: at the 4{,}000~req/s tier, \texttt{pg} achieves
3{,}697 of the offered requests with a corrected p99 of 1.20~s, Prisma achieves
3{,}501 (2.11~s), and Knex achieves 2{,}448 (9.4~s); Sequelize already collapses
at 2{,}000~req/s (achieving 772, with a third of the issued requests timing
out: 9{,}992 of 30{,}000), and
MikroORM collapses at 1{,}000~req/s (achieving 488) and again at 4{,}000
(achieving 259, with 53{,}315 of 60{,}000 requests timing out). We call a layer's behavior at an offered rate a
\emph{collapse} when it achieves under half the offered rate with a majority of
requests timing out. By that criterion the ranking matches the closed-loop result
and sharpens it: the data-mapper ORMs enter unstable queueing with high timeout
rates as soon as the offered load crosses their capacity, whereas the lightweight
layers degrade gradually over a wider range.

\subsection{Measurement stability and significance}
The rankings above rest on stable measurements. Across the 40 PostgreSQL cells
the coefficient of variation of throughput over the 25 independent replicates is
at most $7.6\%$ (\texttt{pg}, insert; Supplement Table~S9), and across the 40
MySQL cells at most $6.6\%$ (Knex, insert; Supplement Table~S1); the
largest relative median absolute deviation is $5.4\%$ (\texttt{pg}, insert,
PostgreSQL). These maxima are well below the between-layer spreads reported
above (up to $6.55\times$). One cell, Knex/PostgreSQL/insert, has $n=24$ rather
than 25, the one write boot in the primary campaign that failed its
post-rebuild health check (Section~\ref{sec:methodology}). A separate order-randomization check (forward versus reversed cell order,
$n=5$ each, insert endpoint) found no detectable history effect
(Section~\ref{sec:methodology}).

To confirm that neighboring ranks are not measurement artifacts, we compare
each adjacently ranked pair on the deep/nested fetch \emph{pairwise} with the
paired permutation test described in Section~\ref{sec:methodology}, reporting
the geometric-mean paired ratio, a paired bootstrap 95\% CI, and the paired
dominance (Table~\ref{tab:significance}); the Wilcoxon signed-rank test agrees
throughout (replication package). Six of the
seven adjacent pairs separate at the permutation-resolution floor
($p \leq 5\times10^{-5}$) with the faster layer ahead in every one of the 25
replicates (dominance $1.00$, except objection\,$>$\,sequelize at $0.88$), all
with ratios at least $1.05$. This is confirmatory analysis over the seven
adjacent-rank comparisons (the confirmatory family), and every one of the six
survives a Bonferroni correction across the family. The seventh, the top pair \texttt{pg} and
Prisma, is the exception treated above: pg leads in 16 of 25 replicates by a
geometric-mean ratio of $1.02$ (95\% CI [1.00--1.04]), which the paired
permutation test does not resolve as significant ($p=0.058$) and the paired TOST
finds equivalent within $\pm5\%$. We therefore read the top of the ranking as a
practical tie, and every step below it as both statistically and practically
supported.

\subsection{Scaling with concurrency}
Figure~\ref{fig:scaling} sweeps the deep/nested fetch from 1 to 256 connections
(a separate sweep on PostgreSQL). The coarse three-band ordering, native and
tuned-native together with Prisma on top, the query builder and the lightweight
ORM in a middle band, and the data-mapper ORMs below, holds at every operating
point at or above 32 connections; at 8 connections Prisma's engine is still
mid-pack (1{,}869~req/s, below Knex's 2{,}438), joining the top band only from
32 upward. Within the top band, \texttt{pg} and
\texttt{pg-tuned} lead Prisma at every concurrency from 8 upward (for example, at
256 connections: \texttt{pg-tuned} 3{,}709, \texttt{pg} 3{,}511, Prisma
3{,}203), so Prisma's near-native parity at the primary 50-connection operating
point does not extend to an outright lead at higher concurrency in this sweep.
At a single connection the ranking differs from the higher-concurrency regime
(the workload is latency- rather than throughput-bound at that point), and
Prisma is among the slowest layers there (503~req/s). The lightweight layers
reach a substantially higher throughput ceiling than the data-mapper ORMs across the
whole load range, so the RQ1 finding is not an artifact of the single
50-connection operating point.

\input{\tabledir fig_scaling}

\subsection{Resource footprint}
Throughput is not the only cost (Supplement Table~S10). Prisma's
near-native or leading throughput on several patterns comes at a much higher CPU
cost: measured at the \emph{same} ten-connection pool as every other layer, its
Node.js process draws $498\%$ (about five cores) on the deep fetch against the
$\approx100\%$ (one core) of the other layers, because Prisma runs a
multi-threaded query engine in-process (Section~\ref{sec:methodology}), a
property of the layer rather than of a larger connection pool; the multiplier
is pattern-dependent (Prisma's aggregation cell draws only $\approx200\%$).

Normalizing throughput by CPU cost makes the trade explicit, but the normalization
must be read at the right tier. Counting only \emph{application}-process CPU on the
deep fetch, requests per CPU-second are \texttt{pg} 3{,}741 down to Prisma 740 and
MikroORM 571, so Prisma ranks second-to-last. That figure understates the compute
of layers that push work to the database, however: \texttt{pg}'s hand-written join
drives about 3.6 database cores while Prisma's engine keeps more work in the
application tier (Supplement Table~S5). Counting \emph{combined} application-plus-database
CPU per completed request, \texttt{pg}'s efficiency advantage over Prisma shrinks
from about $5\times$ (3{,}741 vs.\ 740~req per application-CPU-second) to about
$1.4\times$ (813 vs.\ 596~req per combined-CPU-second); Prisma remains among the
least efficient layers end-to-end, but the gap is a small multiple, not a large
one. The application-tier reading is the one that matters when the application
container's CPU is the budgeted resource; the combined reading is the end-to-end
compute view. An equal-CPU control corroborates the application-tier trade
operationally: confining the server process with \texttt{taskset} to 1, 2, or 4
cores (database and load generator pinned to disjoint cores; Supplement
Table~S4), \texttt{pg} reaches its full deep-fetch throughput on a single
core (3{,}663, 3{,}277, and 3{,}700~req/s at 1, 2, and 4 cores), whereas Prisma
needs roughly four cores to approach that level (896, 1{,}726, 3{,}025),
$4.1\times$ slower than \texttt{pg} on an equal one-core application budget;
MikroORM stays low regardless of core count (463, 577, 542).

Memory differences are smaller. The native driver, Knex, and Objection are
lightest (204--211~MB peak RSS), while the remaining layers use more
(248--271~MB). Resource cost, like throughput, is pattern- and layer-specific,
and matters when sizing horizontally scaled deployments, where per-instance
resource use governs how many replicas a given load
requires~\cite{vaquero2011scaling}.

\input{\tabledir significance_deep_fetch}
\input{\tabledir fig_cpu_tradeoff}
