\section{Threats to Validity}
\label{sec:threats}

We organize the threats along the four standard validity
categories~\cite{cook1979quasi,wohlin2012experimentation}.

\paragraph{Construct validity.}
Our dependent variables are HTTP-level throughput (req/s) and tail latency
(p99) measured at the Express endpoint, not query-execution time at the
driver or database boundary. This is deliberate, since it captures the full request
round trip a deployed service incurs, including Express routing and JSON
serialization, but it means the overhead we attribute to the \emph{access layer}
may partly reflect these surrounding costs instead. We control for this by holding
those costs constant across conditions: every adapter satisfies a documented
\emph{adapter contract} (Section~\ref{sec:methodology}) that guarantees an
identical result shape to Express, so the JSON payload (and its serialization
cost) is the same regardless of which layer produced it. Observed differences
between layers are therefore attributable to how each layer reaches that shape,
not to what is returned. One caveat is specific to writes: durability is relaxed
asymmetrically across the two engines (Section~\ref{sec:methodology}), so absolute
insert throughput is not comparable between PostgreSQL and MySQL. We therefore state
the write finding only within each engine (a $1.3\times$ layer spread on MySQL versus
$4.7\times$ on PostgreSQL), not as a cross-engine magnitude comparison.

\paragraph{Internal validity.}
A naively configured ORM can trigger N+1 queries and appear artificially slow for
reasons unrelated to its inherent cost; every adapter is required to use its
layer's documented, idiomatic eager-loading strategy on the deep-fetch endpoint
(Section~\ref{sec:methodology}). An automated correctness cross-check asserts that
all nine adapters return identical, normalized results (compared as a projection of
key fields on four of the five access patterns) before any timing run is trusted.
This check caught two defects during development: a fan-out bug that inflated the
\texttt{SUM(views)} aggregate for every join-based adapter, and a result-shape
parsing defect in one driver binding, both fixed before the measurements reported here. Explicit warm-up phases
per (layer, engine, endpoint) cell absorb JIT, pool-fill, and plan-cache effects
before timing begins (controls of the kind that guard against environmentally
induced measurement bias~\cite{mytkowicz2009producing}), and two further controls
(Section~\ref{sec:methodology}) rule out non-layer confounds: benchmark-inserted rows are reset before every cell, and the
aggregation endpoint uses one correlated-subquery plan across layers, so it measures
access-layer overhead rather than differences in SQL shape. The write endpoint is
reset to the seeded table before every measured run (not merely before each cell), so
no run inserts into a table grown by an earlier run. A residual risk is coordinated
omission in the load generator, which can
under-report tail latency at saturation~\cite{fruth2022tail,tene2015latency}; a
fixed-rate, constant-arrival (open-loop) cross-run with a coordinated-omission-corrected
generator such as k6 or wrk2 is planned to bound it.

\paragraph{External validity.}
All measurements come from a single schema, a single dataset size, one hardware
configuration, and a connection pool fixed at ten, against specific engine and
library versions (PostgreSQL~18.4, MySQL~9.7.1; Section~\ref{sec:methodology});
results may not transfer to larger working sets, other pool sizes, or other engine
versions, and library performance ages quickly. We do not claim the absolute req/s
figures generalize beyond our hardware; only the \emph{relative} ordering within an
engine is intended to travel. We mitigate this by pinning every dependency
version, publishing the deterministic seed and an automated environment capture,
and releasing the full harness so the matrix can be re-run on other hardware or
against newer releases. A further limitation is that the load generator and server
under test share a host, which is common practice in this genre and removes network
variance as a confound, though it can understate network-bound effects. A
two-machine cross-run over a dedicated link is planned to bound this.

\paragraph{Conclusion validity.}
Our analysis (Section~\ref{sec:methodology}) reports medians with the coefficient of
variation and bootstrap 95\% confidence intervals as stability signals and tests
adjacently ranked layers with the Mann--Whitney~$U$ test and Cliff's~$\delta$
(Tables~\ref{tab:cv}, \ref{tab:significance}). Two bounds reinforce the ordering: the
widest spreads, such as $6.3\times$ on the deep/nested fetch, dwarf the run-to-run CV
(at most 6.9\% on PostgreSQL, Table~\ref{tab:cv}, and 11.2\% on MySQL,
Table~\ref{tab:cv_mysql}), and the ranking is preserved across the full 1--256
connection sweep (Fig.~\ref{fig:scaling}). On the deep/nested fetch every adjacent
pair separates with a maximal effect (Cliff's~$\delta = 1.00$) and non-overlapping
bootstrap CIs, \emph{except} native \texttt{pg} and Prisma at the top, which are
statistically indistinguishable; we read those two as tied rather than ranked. The
Mann--Whitney $p$ is $0.012$ for the separating pairs, near the floor achievable with
five samples per group, so with only five replicates a Bonferroni correction across
the pairwise tests is not passed on $p$ alone; the maximal effect sizes and
non-overlapping intervals carry the inference. Crucially, because each replicate is a
freshly booted server process with cell order randomized, these are genuine
\emph{independent} replicates rather than repeated measures within one process
lifetime~\cite{kalibera2013rigorous}; more replicates or longer runs would tighten
the intervals further.
