\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amssymb,amsmath}
\usepackage[protrusion=true,expansion=false]{microtype}
\usepackage{enumitem}
\setlist{nosep,leftmargin=1.5em}
\usepackage{parskip}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,urlcolor=blue}
\newcommand{\euro}{EUR}
\pagestyle{plain}
\begin{document}
Gdańsk, Poland, 14 July 2026

Dear Editors,

Please consider my manuscript, "A Reproducible Benchmark of Relational Database Access-Layer Performance in Express.js: A Configuration-Specific Comparison on PostgreSQL and MySQL," for publication in \emph{Information and Software Technology} as a Research Paper.

Node.js/Express services reach relational databases through a spectrum of access layers: native drivers, query builders, and object-relational mappers (ORMs). This choice is made on nearly every non-trivial back-end project, yet in practice it is guided by vendor benchmarks that are fragmented: each is published by the maintainers of one of the compared products, almost all target PostgreSQL exclusively, and each reports a single metric family (throughput or latency, rarely the p99 tail that dominates perceived performance under load). The peer-reviewed literature is sparser still: the closest prior study compared ORM overhead across eight frameworks but explicitly excluded JavaScript, and two 2025 studies of Node.js/TypeScript access layers are each limited to PostgreSQL and to at most three ORMs. In the sources I searched through July 2026, I did not identify a study combining broad taxonomy coverage, both major open-source engines, vendor independence, and joint throughput/tail reporting.

The paper's primary contribution is a reproducible, byte-equivalence-checked benchmark instrument, which I use for a configuration-specific comparison rather than a generalized library ranking. An identical Express application is served through nine idiomatic access layers (the native drivers pg and mysql2, the query builder Knex, and six ORMs: Drizzle, Prisma, Sequelize, TypeORM, Objection, and MikroORM) plus two tuned native baselines, over five access patterns, against both PostgreSQL and MySQL, reporting throughput and saturated client-observed response-time p99 jointly across all 90 measured configurations. Because within each replicate every layer is driven by the identical request stream, the design is a randomized block, and layers are compared with paired tests on their per-replicate ratios; each configuration is measured over 25 repeated runs (a freshly booted server process per run, cell order randomized), and before any measurement an automated cross-check verifies that every layer returns byte-identical responses. Inserts are measured under each engine's default durability. A same-SQL control bounds the largest effect observed, a native-relative spread of 6.5x between the native driver and the slowest data-mapper ORM on a deep, nested fetch: on identical SQL (byte-identical statements confirmed from server-side logs, wire protocols disclosed) that spread narrows to at most 1.5x, but because the control changes query formulation, protocol, statement preparation, round-trips, and mapping together, it bounds the residual difference without isolating any single factor. A utilization-controlled open-loop experiment shows the saturated p99 gap largely dissolves at matched utilization, and a multi-worker experiment shows the native driver overtakes the query-engine ORM once every layer uses the otherwise-idle cores---so several apparent effects are shown to be configuration artifacts rather than intrinsic properties. Further controls (equal-CPU and combined application-plus-database CPU accounting, a per-layer pool-size frontier, a mixed read/write workload, and durability, fan-out, and post-restart checks) accompany these. The complete replication package is public on GitHub and permanently archived on Zenodo (DOI: 10.5281/zenodo.21427844); a numbered online supplement accompanies the manuscript.

The manuscript fits \emph{Information and Software Technology} directly: it is a controlled empirical study of a decision that essentially every Node.js/Express project makes---which access layer to adopt---aimed at improving that practice with evidence rather than vendor marketing. It is reported to the standard the journal expects of empirical software engineering work: controlled factors, repeated randomized-block runs, paired nonparametric significance and equivalence testing with effect sizes and per-cell confidence intervals, an explicit threats-to-validity analysis, and an open, reusable replication package.

The work is original, has not been published previously, and is not under consideration elsewhere. I declare no competing interests and received no funding for this research. A declaration of generative-AI use in the manuscript-preparation process is included in the manuscript in accordance with Elsevier policy.

I look forward to your editorial decision.

Sincerely,

Mateusz Miotk Faculty of Mathematics, Physics and Informatics, University of Gdańsk, Poland mateusz.miotk@ug.edu.pl (corresponding author)

\end{document}
