%% Information and Software Technology (Elsevier) submission version.
%% Class: elsarticle, citation style elsarticle-num (per the IST guide for authors).
%% Build: latexmk -pdf ist_main.tex   (bibtex; inputs ../sections and ../tables).
%% `review` gives single-column, double-spaced, line-numbered output for reviewers.
\documentclass[review]{elsarticle}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{array}
\usepackage{adjustbox}
\usepackage{enumitem}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[protrusion=true,expansion=false]{microtype}
\setlength{\emergencystretch}{3em}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{url}

% Shared sections resolve result tables through \tabledir; from paper/ist/ the
% tables live one level up.
\newcommand{\tabledir}{../tables/}

\journal{Information and Software Technology}

\begin{document}

\begin{frontmatter}

\title{A Comparability Protocol for Benchmarking Relational Database
  Access Layers in Express.js}

\author[ug]{Mateusz Miotk\corref{cor1}}
\ead{mateusz.miotk@ug.edu.pl}
\cortext[cor1]{Corresponding author.}
\affiliation[ug]{organization={Faculty of Mathematics, Physics and Informatics,
    University of Gda\'nsk},
  addressline={Wita Stwosza 57},
  postcode={80-308},
  city={Gda\'nsk},
  country={Poland}}

\begin{abstract}
\textbf{Context:} Node.js/Express services reach relational databases through a
spectrum of access layers (native drivers, query builders, and object-relational
mappers, or ORMs), yet practitioners choose largely on vendor
benchmarks whose numbers are rarely shown to be comparable --- fragmented,
PostgreSQL-only, and seldom reporting the tail (p99).
\textbf{Objective:} We contribute a reusable \emph{comparability protocol} for
access-layer benchmarking --- a semantic-equivalence correctness gate, a
documentation-selected estimand paired with a same-SQL standardized contrast, and separated
operating points --- realized in an open harness and demonstrated through a
dual-engine PostgreSQL/MySQL case study.
\textbf{Methods:} An identical Express application is served through nine
documentation-selected access layers and two tuned native baselines, over five access
patterns, on the latest stable versions (July 2026 freeze).
We report throughput with the \emph{high-load response-time p99 under equal closed-loop demand}
(a fixed 50-connection point) from 25 repeated runs per cell, and on the deep fetch distinguish it
from the tail at \emph{equal utilization}, where the gap converges --- the near-saturation p99
chiefly a capacity-and-queueing outcome, not a sub-saturation latency difference.
An automated cross-check verifies byte-identical responses before measurement.
\textbf{Results:} The implementation-and-strategy difference is largest on the tested deep/nested fetch: spreads $7.15\times$
(PostgreSQL) and $4.96\times$ (MySQL) between native and the slowest ORM,
against only $1.68\times$ and $2.01\times$ among the raw paths running common SQL --- a
standardized contrast that isolates no single mechanism. The read ordering holds
across engines, but the aggregation and insert orderings
reverse: Prisma drops to last on the MySQL insert. No layer exceeds
$\approx\!110\%$ of one core; the native driver leads throughout.
\textbf{Conclusion:} Access-layer
rankings are configuration-specific and version-sensitive, so the durable
contribution is the protocol and its harness, not any single ranking.
\end{abstract}

\begin{keyword}
benchmarking methodology \sep database access layer \sep object-relational mapping
\sep Node.js \sep response-time tail (p99) \sep empirical software engineering
\end{keyword}

\end{frontmatter}

%% ------------------------------------------------------------------
\input{../sections/introduction}
\input{../sections/related_work}
\input{../sections/methodology}
\input{../sections/results}
\input{../sections/discussion}
\input{../sections/threats}
\input{../sections/conclusion}

%% ------------------------------------------------------------------
\section*{CRediT authorship contribution statement}
\textbf{Mateusz Miotk:} Conceptualization, Methodology, Software, Investigation,
Data curation, Formal analysis, Visualization, Writing -- original draft,
Writing -- review \& editing.

\section*{Declaration of competing interest}
The author declares that he has no known competing financial interests or personal
relationships that could have appeared to influence the work reported in this paper.
The author is not affiliated with, nor funded by, any of the database libraries or
vendors evaluated in this study.

\section*{Data availability}
The complete replication package (benchmark harness, database schema and
deterministic seed, all adapter implementations, the byte-level correctness cross-check,
raw per-cell measurements, and the scripts that generate every table in this
paper) is publicly available at
\url{https://github.com/mmiotk/express-db-access-performance} and permanently
archived at Zenodo as release v1.12.12
(\url{https://doi.org/10.5281/zenodo.21498786}). A reproducibility summary --- versions,
requirements, the run commands, the time and resources needed, and which results the
harness regenerates automatically from the archived raw data --- is in Supplement
Table~S31 and \texttt{REPRODUCE.md}. \textbf{Exact reproduction requires the reference engines}
(PostgreSQL~18.4 and MySQL~9.7.1), installed via the user-space conda path documented in
\texttt{REPRODUCE.md}; the convenience Docker path pins \emph{older} engines (PostgreSQL~16,
MySQL~8.4) and reproduces the workflow but not the published numbers. A machine-readable encoding of
the protocol --- inputs, mandatory and recommended stages, the cell-admission gate, and compliance
levels --- ships as \texttt{protocol-checklist.yaml} for auditing a benchmark against the protocol. A clean-room reproduction from the archived tarball (35/35 raw-data checksums; byte-for-byte regeneration of 45 of 50 tables) is logged in \texttt{notes/clean-room-reproduction.md}.

\section*{Funding}
This research did not receive any specific grant from funding agencies in the
public, commercial, or not-for-profit sectors.

\section*{Declaration of generative AI and AI-assisted technologies in the writing process}
During the preparation of this work the author used a large language model (Claude,
Anthropic) in order to draft and edit the manuscript text and improve its readability
and language. After using this tool, the author reviewed and edited the content as needed
and takes full responsibility for the content of the published article.

\section*{AI-assisted research software}
Separately from the writing above, a large language model (Claude, Anthropic) also assisted
the author in implementing and analyzing the benchmark harness. This is research tooling for
which the author takes full responsibility, and all AI-assisted analytical code is
independently inspectable and verified: the statistical estimators carry unit tests against
hand-computed values and known closed-form properties (\texttt{bench/stats.test.mjs}, run by
\texttt{npm test}), every reported table cell traces to a raw per-cell record through a
committed generator (\texttt{MANIFEST.md}), and the byte-level correctness cross-check
independently confirms that all layers return identical responses before any timing is
trusted.

%% ------------------------------------------------------------------
\bibliographystyle{elsarticle-num}
\bibliography{../references}

\end{document}
