\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amssymb,amsmath}
\usepackage[protrusion=true,expansion=false]{microtype}
\usepackage{enumitem}
\setlist{nosep,leftmargin=1.5em}
\usepackage{parskip}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,urlcolor=blue}
\newcommand{\euro}{EUR}
\pagestyle{plain}
\begin{document}
\section*{Suggested reviewers --- IST submission (verified 2026-07-11)}

All affiliations/emails verified against official homepages (emails quoted exactly as published; none guessed). CoI screen: none holds an IST editorial role (EiC: Günther Ruhe; Special Content Editor: Jeffrey Carver); none has a Poland/Gdańsk link. Before submitting, glance at the IST editorial-board page once in a browser: https://www.sciencedirect.com/journal/information-and-software-technology/about/editorial-board

Pick 3--5 for the Editorial Manager form, in this order:

\begin{enumerate}
  \item \textbf{Tse-Hsun (Peter) Chen} --- Associate Professor, Computer Science \& Software
\end{enumerate}

Engineering, Concordia University, Montreal, Canada. peterc@encs.concordia.ca --- https://petertsehsun.github.io/ Rationale: author of the ORM performance anti-pattern line (ICSE'14, TSE'16, FSE'16) that this paper directly extends; still active on database-access performance (TOSEM 2024, ICSE/ASE 2025).

\begin{enumerate}
  \item \textbf{Stefanie Scherzinger} --- Full Professor, Scalable Database Systems, University
\end{enumerate}

of Passau, Germany. Stefanie.Scherzinger@uni-passau.de --- https://www.fim.uni-passau.de/en/database-systems Rationale: co-author of "Tell-Tale Tail Latencies" (TPCTC'21) --- the tail-latency and DB-benchmarking-pitfalls methodology at the core of this paper; reproducibility-focused.

\begin{enumerate}
  \item \textbf{Philipp Leitner} --- Associate Professor (Docent), Chalmers | University of
\end{enumerate}

Gothenburg, Sweden. philipp.leitner@chalmers.se --- https://www.philippleitner.net/ Rationale: leading expert on software microbenchmarking rigor (EMSE'19, ICPE): replicates, variability, effect sizes --- this paper's methodological frame. (AE of EMSE --- no conflict for IST.)

\begin{enumerate}
  \item \textbf{Romain Rouvoy} --- Full Professor, Université de Lille / CRIStAL (CNRS) / Inria,
\end{enumerate}

France. romain.rouvoy@univ-lille.fr --- https://pro.univ-lille.fr/romain-rouvoy Rationale: SANER 2024 ORM performance/energy study --- the most recent ORM-benchmarking work among the paper's references.

\begin{enumerate}
  \item \textbf{Alvin Cheung} --- Full Professor, EECS, UC Berkeley, USA. (No published email;
\end{enumerate}

contact via homepage.) https://people.eecs.berkeley.edu/\textasciitilde{}akcheung/ Rationale: foundational work on optimizing database-backed applications (query synthesis, PLDI'13 onward); active at the DB/PL intersection.

\begin{enumerate}
  \item \textbf{Cor-Paul Bezemer} --- Associate Professor \& Canada Research Chair, ECE,
\end{enumerate}

University of Alberta, Canada. bezemer@ualberta.ca --- https://www.ece.ualberta.ca/\textasciitilde{}bezemer/ Rationale: performance-engineering state of practice and empirical performance studies (EMSE/TOSEM 2025); ranked last as recent output partly shifted to game/AI QA.

Replacement (if Chen unavailable): \textbf{Weiyi Shang} --- Full Professor, ECE, University of Waterloo, Canada, wshang@uwaterloo.ca (co-author of the same ORM papers as Chen --- do not list both).

Excluded after verification: Guoliang Jin (at Databricks since 2023, research inactive), Christoph Laaber (Leitner's former PhD student --- entangled with pick \#3), Alexandre Torres (no verifiable current affiliation), Renata Galante (current activity in this area unconfirmed; possible 7th option after checking her UFRGS page).

Diversity: five countries in the top 6; at most two from one country; no two co-authors of the same paper in the top 5.

\end{document}
