% Online supplement to:
%   A Comparative Analysis of Relational Database Access-Layer Performance
%   in Express.js for PostgreSQL and MySQL
% Build: latexmk -pdf supplement.tex   (from paper/; tables in tables/)
\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{array}
\usepackage{adjustbox}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage[protrusion=true,expansion=false]{microtype}
\usepackage{url}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, urlcolor=cyan,
  pdftitle={Online supplement: Relational Database Access-Layer Performance in Express.js}}

\newcommand{\tabledir}{tables/}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\title{Online supplement:\\
  A Comparability Protocol for Benchmarking Relational Database\\
  Access Layers in Express.js}
\author{Mateusz Miotk\\
  Faculty of Mathematics, Physics and Informatics, University of Gda\'nsk\\
  \texttt{mateusz.miotk@ug.edu.pl}}
\date{}

\begin{document}
\maketitle

\noindent This supplement carries measurement detail referenced from the main
text. The complete revised artifact is archived as release v1.12.13 under the Zenodo concept DOI (\url{https://doi.org/10.5281/zenodo.21313858}). Its primary-campaign tables are unchanged from release v1.12.12 (version DOI \url{https://doi.org/10.5281/zenodo.21498786}); v1.12.13 additionally carries the documentation archive and standalone canonical-constructor measurement, each with separate manifests and inputs.

\section{Dispersion on MySQL}
\label{sup:cv-mysql}
The main text reports the coefficient-of-variation table for PostgreSQL and
the per-engine maxima. Table~\ref{tab:cv_mysql} gives the full per-layer,
per-pattern CV on MySQL across the 25 repeated runs (maximum 15.9\%,
\texttt{drizzle}/insert; the insert cells are the noisiest in the study).
Figure~\ref{fig:insert_dispersion} plots the raw per-replicate insert throughput
behind these coefficients: several cells are visibly bimodal or heavy-tailed
within a cell, structure a single CV cannot convey, which is why the insert rank
correlation and dispersion are reported conservatively.

\input{\tabledir cv_mysql}

\input{\tabledir fig_insert_dispersion}

\section{Documentation-selected round-trip counts}
\label{sup:query-counts}
Table~\ref{tab:query_counts} reports, for each layer, the number of SQL
statements the documentation-selected deep fetch issues per request, captured from
server-side statement logs. The counts motivate the main text's observation
that round-trip count alone does not determine throughput: the two layers
issuing four statements (Prisma, Objection) sit at opposite ends of the
throughput ladder.

\input{\tabledir query_counts}

\section{Write throughput under default versus relaxed durability}
\label{sup:durability}
Table~\ref{tab:durability} gives the per-layer insert throughput under each
engine's default durability configuration (the paper's primary write regime)
and under the symmetric relaxed regime (secondary, 10 replicates), with the
relaxed$\div$default ratio quoted in the main text.

\input{\tabledir durability}

\section{Equal-CPU-budget sensitivity check}
\label{sup:equalcpu}
Table~\ref{tab:equalcpu} gives the deep-fetch throughput with the server
confined to 1, 2, or 4 cores (database and load generator pinned to disjoint
cores); all nine values are quoted in the main text.

\input{\tabledir equalcpu}

\section{Application-tier versus end-to-end CPU efficiency}
\label{sup:cpu}
Table~\ref{tab:cpu_efficiency} reports, per layer on the deep fetch,
the median application-process-tree CPU and database-server CPU, the combined
core-count, and throughput per application-CPU-second and per combined
(application$+$database) CPU-second. The application-only figure understates the
compute of layers that delegate more work to the database, but on the current
versions the application-only and combined figures largely agree, because no
layer draws more than about one application core: \texttt{mysql2} leads
application-tier efficiency by about $3.8\times$ over Prisma (2{,}269 vs.\
604~req per app-CPU-s) and by $3.3\times$ end-to-end (1{,}254 vs.\ 384~req per
combined-CPU-s). There is no CPU-for-throughput trade of the kind an earlier
version of this study reported.

\input{\tabledir cpu_efficiency}

\section{Coordinated-omission-corrected open-loop sweep}
\label{sup:openloop}
Table~\ref{tab:openloop} gives the full constant-arrival sweep (PostgreSQL,
five layers, offered rates 250--4{,}000 requests/s) behind the open-loop
paragraph of the main text: achieved throughput, corrected latency
percentiles measured from each request's \emph{intended} send time, and
timeout counts (10~s cap, clipped into the distribution).

\input{\tabledir openloop}

\section{Pool-size sensitivity}
\label{sup:poolsize}
The primary campaign fixes every layer's connection pool at 10 to isolate the
access layer from pool tuning. Table~\ref{tab:poolsize} varies the pool from 1 to
50 for a representative layer of each tier on the deep fetch (PostgreSQL, 50
connections). The deep-fetch ordering (\texttt{pg} $>$ Knex $>$ Prisma $>$ MikroORM) is preserved at every pool size, and each layer's
throughput saturates at a pool well below 50, so the fixed-pool choice does not
appear to drive the ranking; a very small pool (1--2) is the one regime that compresses the
differences, because every layer then queues on connection acquisition.

\input{\tabledir poolsize}

\section{Transactional multi-statement write}
\label{sup:txn}
Beyond the single-row insert of the primary matrix, Table~\ref{tab:txn_write}
reports a transactional write, \texttt{POST /threads}, which inserts a post and
five comments in one transaction through each layer's transaction facility, for a
representative layer of each tier on PostgreSQL under default durability (median of five
runs, exploratory). The
full layer\,$\times$\,engine matrix for this pattern is left to future work. Because
the transaction commits once, its per-commit fsync dominates and the layers cluster
more tightly than the widest read spread ($3.7\times$, \texttt{pg} 2{,}718 down to
Prisma 733), though the ordering does not simply track the single-row insert:
Prisma, mid-pack on the single-row PostgreSQL insert, falls to the bottom of the
transactional subset.

\input{\tabledir txn_write}

\section{Dispersion on PostgreSQL}
\label{sup:cv-pg}
Table~\ref{tab:cv} gives the full per-layer, per-pattern coefficient of variation
of throughput on PostgreSQL across the 25 repeated runs (maximum 7.3\%,
\texttt{knex}/insert), the companion to the MySQL table in Section~S1.

\input{\tabledir cv_all}

\section{Resource footprint}
\label{sup:resources}
Table~\ref{tab:resources} gives the per-layer application-process CPU (median \%
of one logical core) and peak resident memory on the deep fetch
(MySQL), the detail behind the CPU trade-off discussed in the main text
(all layers $\approx100$--$110\%$ of one core; peak RSS 215--356~MB).

\input{\tabledir resources}

\section{Pinned versions}
\label{sup:versions}
Table~\ref{tab:versions} lists the pinned versions of every evaluated access layer
and tool, recorded from the lockfile on the measurement date.

\begin{table}[htbp]
  \centering
  \caption{Pinned versions of the evaluated access layers and tooling.}
  \label{tab:versions}
  \begin{tabular}{l l l l}
    \toprule
    Layer / tool & Version & Layer / tool & Version \\
    \midrule
    \texttt{pg}         & 8.22.0  & \texttt{sequelize}      & 6.37.8 \\
    \texttt{mysql2}     & 3.23.0  & \texttt{typeorm}        & 1.1.0  \\
    \texttt{knex}       & 3.3.0   & \texttt{objection}      & 3.1.5  \\
    \texttt{drizzle-orm}& 0.45.2  & \texttt{@mikro-orm/core}& 7.1.6  \\
    \texttt{@prisma/client} & 7.8.0 & \texttt{express}     & 5.2.1  \\
    \texttt{autocannon} & 8.0.0   & Node.js                 & 24.18.0 \\
    PostgreSQL          & 18.4    & MySQL                   & 9.7.1  \\
    \bottomrule
  \end{tabular}

  \vspace{2pt}
  \footnotesize Pinned to the latest stable release compatible with the harness
  adapter contract at the freeze date (2026-07-15); Sequelize's version-7 line is
  a prerelease, so the 6.x stable line is used. Prisma~7 is Rust-free and connects
  through an official JS driver adapter (\texttt{@prisma/adapter-pg} for PostgreSQL,
  \texttt{@prisma/adapter-mariadb} with the \texttt{mariadb} 3.5.3 driver for MySQL,
  as Prisma ships no \texttt{mysql2} adapter).
\end{table}

\section{Per-pattern detail: the read patterns}
\label{sup:reads}
The two cheapest read patterns show the smallest between-layer spread and are
summarized in the main text; their full throughput and p99 tables are given here.
Table~\ref{tab:point_read} reports the point read (primary-key lookup) and
Table~\ref{tab:range_scan} the keyset range scan, each by access layer and engine.
The consequential patterns (deep fetch, aggregation, insert) are tabulated
in the main text.

\input{\tabledir point_read}
\input{\tabledir range_scan}

\section{Same-SQL standardized contrast: full table}
\label{sup:sameplan}
Table~\ref{tab:sameplan} gives the full same-SQL (raw-path) standardized contrast on the deep
fetch for both engines: throughput when every layer executes the identical
statement (verified byte-identical by server-side capture). The main text
reports the resulting smaller residual spread as a standardized-SQL contrast on the
residual same-SQL difference --- a compound intervention, not a bound, that isolates no single factor.

\input{\tabledir sameplan}

\section{Per-pattern detail: aggregation}
\label{sup:aggregation}
Table~\ref{tab:aggregation} gives the full aggregation-pattern throughput and p99
by access layer and engine. RQ3 finds aggregation the least layer-sensitive
pattern once every layer runs the same raw correlated-subquery plan; the main text
reports the ranking and the leading figures.

\input{\tabledir aggregation}

\section{Documentation-selected deep-fetch choices per access layer}
\label{sup:adapter-choices}
Table~\ref{tab:adapter_choices} documents, for each access layer, the exact
eager-loading API used on the deep fetch, the number of SQL round trips it
issues, the documented alternative loading strategy we did \emph{not} use, and the
query-logging state. Every layer uses its documented default path;
query logging is disabled and verified on all eleven adapters, and no lifecycle
hooks or validation run on the read path. The six data-mapper and ORM layers split
both ways on the join-versus-select-in default: Sequelize, TypeORM, and MikroORM
default to a single join, while Prisma and Objection default to separate batched
queries. The loading-strategy sensitivity check reported in the main text
(Table~\ref{tab:altloading}) switches the join-default layers to select-in where
that alternative is a byte-identical drop-in. Full per-adapter detail is in the
replication package's \texttt{METHODOLOGY.md}.

\input{\tabledir adapter_choices}

\section{Open-loop tail latency on MySQL}
\label{sup:openloop-mysql}
Table~\ref{tab:openloop_mysql} is the MySQL companion to the PostgreSQL open-loop
sweep (Supplement Table~S6): coordinated-omission-corrected p99 on the deep
fetch at a constant offered rate. The sub-saturation tail is flat and small on
both engines, and each layer collapses once the offered rate crosses its capacity,
so the high-load-tail ordering holds on MySQL as on PostgreSQL.

\input{\tabledir openloop_mysql}

\section{Alternative eager-loading sensitivity}
\label{sup:altloading}
Table~\ref{tab:altloading} reports the loading-strategy sensitivity check: for the
data-mapper ORMs whose alternative strategy is a byte-identical drop-in, documentation-selected
deep-fetch throughput versus the alternative, on both engines. Both endpoints were
verified to return byte-identical responses before timing.

\input{\tabledir altloading}

\section{MySQL insert commit-flush mechanism}
\label{sup:waitevents}
Table~\ref{tab:waitevents} gives the direct \texttt{performance\_schema} evidence
behind the MySQL insert bottleneck: sampling the redo-log and binary-log flush wait
instruments around a fixed insert workload at default durability, the flush is a
shared engine floor---similar per-insert across the native driver, query builder,
and ORM despite their very different read performance---and relaxing durability
removes it.

\input{\tabledir waitevents}

\section{Post-restart environmental-robustness check}
\label{sup:postreboot}
Table~\ref{tab:postreboot} reports the environmental-robustness check: as the
campaign's final stage, after a full restart of both database engines (cold buffer
pools, fresh connections), a representative deep-fetch subset was re-measured against
the primary median. The layer ranking reproduces on both engines
(\texttt{pg}~$>$~Prisma~$\gg$~MikroORM on PostgreSQL; \texttt{mysql2}~$\gg$~Prisma~$>$~MikroORM
on MySQL); absolute throughput is up to $16\%$ lower, a cold-buffer and run-to-run
drift the paper's relative analysis is designed to absorb. Because PostgreSQL and
MySQL cells are not measured simultaneously, this drift is non-uniform across engines
($0\%$ on \texttt{mysql2}, $16\%$ on \texttt{pg}, $16\%$ on MikroORM/PostgreSQL), so
the cross-engine interaction ratios (Table~\ref{tab:interaction}) could in principle
carry some residual drift; the cross-stack differences there are far larger than the $16\%$
drift, so the backend-stack-transfer conclusion holds.

\input{\tabledir postreboot}

\section{Utilization-controlled tail latency}
\label{sup:utilization}
Tables~\ref{tab:utilization} and~\ref{tab:utilization_mysql} give the
coordinated-omission-corrected p99 on the deep fetch when each layer is offered
50/70/85/95\% of its $\widehat C_{50}$ capacity proxy (the 25-run median
50-connection deep-fetch rate), replicated five times, on both engines. At nominally matched utilization the
tails are small and similar across layers---for example on PostgreSQL every layer
holds p99 near 2--4~ms at 50\%---so the wide high-load p99 ladder in the main text is
a capacity-and-queueing effect, not a difference in latency at a matched offered-load fraction; the tail rises only as a layer nears its own capacity. Re-expressing the loads against the
full-sweep maximum leaves the 50\% and 70\% bands below saturation (Table~\ref{tab:capacity_sensitivity}). This is the
capacity-normalized latency
companion to the high-load primary p99. The 50\% and most 70\% cells, which carry that
comparison, are stable across runs; the 85/95\% cells are increasingly noisy (p99
varies several-fold between runs as a layer approaches saturation), so those columns
are read as trend, not precise values.

\input{\tabledir utilization}
\input{\tabledir utilization_mysql}

\section{Longer-window tail sensitivity}
\label{sup:taillong}
The primary p99 is measured over a 12~s window, which yields only $\approx60$
requests above the p99 threshold in the slowest deep-fetch cells. To check that this
per-run tail count is sufficient, Table~\ref{tab:taillong} re-measures the deep fetch
over a 60~s window (about five times the requests, $\approx300$ tail observations per
run) at the same 50-connection operating point, ten repeated runs, on both engines.
The p99 ranking is preserved exactly (Spearman $\rho=1.00$ per engine), each cell's p99
barely moves from the 12~s value, the p97.5 ordering matches the p99 ordering
($\rho=1.00$), and the run-to-run p99 coefficient of variation stays small (maximum
6.0\% on PostgreSQL, 3.8\% on MySQL). The shorter window is therefore adequate for the
tail \emph{ranking}, and the tail conclusions do not rest on the exact per-run tail
count.

\input{\tabledir taillong}

\section{Multi-worker scaling}
\label{sup:cluster}
Table~\ref{tab:cluster} scales a representative subset (three layers per engine) to 1, 2, and 4
Express workers (a shared-port \texttt{node} cluster), median of three runs. Because each process
uses about one core and gains nothing from more, aggregate throughput rises roughly in proportion
to workers over this range: on PostgreSQL the native driver leads at one worker
(\texttt{pg} 3{,}299 to Prisma's 1{,}117); its aggregate throughput then plateaus by two
workers (4{,}055, then 3{,}961 at four) --- whether from database contention, shared-host CPU
competition, or the fixed-connection generator is not instrumented here --- while Prisma, low per
process, rises 1{,}117 to 2{,}275 to 4{,}449 and reaches native-like aggregate
throughput at four workers. This is a bounded check, not a demonstration of general scalability:
it stops at four workers, measures no database-side saturation (DB CPU, connections, or wait
events), and on MySQL horizontal scaling \emph{widens} the gap --- the native \texttt{mysql2}
keeps scaling to 8{,}883~req/s at four workers while Prisma reaches only 2{,}427 (about
\texttt{mysql2} at one worker). Within these limits a layer's low per-process throughput may be
addressed by additional application processes until another bottleneck is reached, at a
proportional cost in cores --- a partial test of the single-host resource-competition concern.

\input{\tabledir cluster}

\section{Pool-size frontier on MySQL}
\label{sup:poolsize-mysql}
Table~\ref{tab:poolsize_mysql} is the MySQL companion to the PostgreSQL pool sweep
(Supplement Table~S7): deep-fetch throughput as the connection pool varies from 1 to
50, per layer. Each layer saturates at a pool well below 50 and the ordering is
preserved at every pool size, so the fixed pool of ten does not appear to drive the ranking on
either engine.

\input{\tabledir poolsize_mysql}

\section{Mixed read/write workload}
\label{sup:mixed}
Table~\ref{tab:mixed} interleaves the deep-fetch read with single-row inserts at
90:10 and 70:30 read:write, both engines, with a physical write reset per run. The
layer ordering matches the read ordering and is insensitive to the mix, because the
deep-fetch read dominates the cost; exploratory.

\input{\tabledir mixed}

\section{Concurrency sweep}
\label{sup:scaling}
Figure~\ref{fig:scaling} sweeps the deep fetch from 1 to 256 client
connections for a representative layer of each tier, both engines. The 50-connection
operating point used throughout sits above the knee for every layer on this
pattern; the ordering is load-robust above about 32 connections. This is the evidence behind the choice of
operating point.

\input{\tabledir fig_scaling}

\section{Per-layer cross-engine ranks}
\label{sup:ranks}
Table~\ref{tab:ranks} gives each portable layer's throughput rank on PostgreSQL
versus MySQL for the deep fetch, aggregation, and the insert, showing the concrete
rank reversals behind the main text's backend-stack-transfer finding. The deep-fetch
orderings largely agree, but aggregation and the insert reverse at the top: on
aggregation \texttt{drizzle} (PostgreSQL rank 1) and \texttt{knex} (MySQL rank 1)
trade the lead, and on the insert \texttt{knex} drops from rank 1 to 3 while
\texttt{drizzle} rises to 1, with Prisma falling from mid-pack (rank 4) to last
(rank 7) on both. The rank correlations these reversals produce ($\rho=0.86$ deep
fetch, $0.64$ aggregation, $0.68$ insert; coarse at $n=7$) are reported only as
descriptive summaries.

\input{\tabledir ranks}

Because MikroORM is the largest single outlier in several spreads, we recompute the
cross-engine agreement without it: over the remaining six portable layers the read
ordering stays similar (deep-fetch $\rho=0.77$, $\tau=0.60$; range $\rho=0.83$,
$\tau=0.73$) and the deep-fetch native-relative spread stays substantial at
$3.72\times$ (PostgreSQL) and $3.76\times$ (MySQL), so the RQ2 and RQ3 findings do not depend on that one
layer.

\section{Layer-by-engine interaction magnitude}
\label{sup:interaction}
The blocked permutation test reports \emph{that} the layer ordering interacts with the
engine; Table~\ref{tab:interaction} reports \emph{how large} the interaction is, as the
PostgreSQL$\div$MySQL throughput ratio per layer and pattern with a 95\% paired-bootstrap
interval. Every layer is faster on PostgreSQL, so the interaction is in the spread of
the backend-stack ratios: the reads stay within $\approx$1.0--1.9 while the
insert scatters from $1.6$ to $3.2$, the quantitative counterpart of the low insert rank
correlation.

\input{\tabledir interaction}

\section{Study factors}
\label{sup:factors}
Table~\ref{tab:factors} records which factors the benchmark holds constant, which
it varies, and which are uncontrolled on the single virtualized host; the
uncontrolled factors are why results are reported for this configuration rather
than as a general library ranking.
\input{\tabledir factors}

\section{Paired significance of the p99 ladder}
\label{sup:significance-p99}
Table~\ref{tab:significance_p99} is the p99 counterpart of the main-text throughput
significance table (Table~6): the same paired permutation, Wilcoxon,
geometric-mean-ratio, paired-bootstrap-CI, and dominance analysis applied to each
adjacent layer's per-replicate deep-fetch p99 on both engines, delivering the per-cell
p99 inference the outcomes table (Table~3) lists as a primary analysis. Every adjacent pair separates on both engines except two near-ties at ratio
$1.03$ --- the TypeORM--Prisma step on PostgreSQL (permutation $p=0.05$) and the
Sequelize--Objection step on MySQL ($p=0.08$); the p99 ordering otherwise mirrors
the throughput ranking. Because the high-load tail tracks capacity,
these are resolved capacity-and-queueing differences, not differences in sub-saturation latency.

\input{\tabledir significance_p99}

\section{Extended methods (in the replication package)}
The reporting checklist, the benchmarking-pitfalls checklist, the detailed statistical methods, and
the measurement details are distributed with the replication package
(\texttt{notes/supplement-methods.pdf}) rather than reproduced here, to keep this supplement compact;
the numbered tables and figures above are unaffected.

\section{Artifact reproducibility summary}
\label{sup:reproducibility}
Table~\ref{tab:reproducibility} summarizes what is needed to reproduce the study and which
results are regenerated automatically; \texttt{REPRODUCE.md} in the replication package is
the single runnable entry point.

\begin{table}[htbp]
  \centering
  \small
  \caption{Artifact reproducibility summary: version and identifier, requirements, how to
    run, time and resources, and which results are reproduced automatically.}
  \label{tab:reproducibility}
  \begin{tabular}{@{}p{0.24\textwidth} p{0.68\textwidth}@{}}
    \toprule
    Item & Detail \\
    \midrule
    Artifact \& version & Complete revision: release v1.12.13 (git tag \texttt{v1.12.13}), including the documentation snapshots, canonical-constructor measurement, and capacity-denominator analysis. The primary measurement matrix is unchanged from v1.12.12. \\
    Persistent identifier & Zenodo concept DOI \texttt{10.5281/zenodo.21313858}, resolving to release v1.12.13; primary-campaign v1.12.12 version DOI \texttt{10.5281/zenodo.21498786}. \\
    Software & Node.js 24.18.0, PostgreSQL 18.4, MySQL 9.7.1, \texttt{npm}. Reference path:
      conda user-space (\texttt{conda-forge}); the Docker path is a convenience and pins
      older engines (PostgreSQL 16 / MySQL 8.4), so it only approximates the published
      numbers. \\
    Hardware \& disk & Single Linux host (full host and version capture in
      Table~\ref{tab:versions}); $\approx$1~GB for the seeded databases. \\
    Deterministic seed & 2{,}000 authors / 100{,}000 posts / 1{,}000{,}000 comments from a
      fixed PRNG. \\
    Smoke test ($\approx$5--10~min) & \texttt{npm ci}; \texttt{npm run migrate \&\& npm run
      seed}; \texttt{npm run bench:quick}; \texttt{node bench/verify.mjs} (read byte-equivalence,
      must print ALL BYTE-IDENTICAL) and \texttt{node bench/verify-writes.mjs} (write-state
      correctness, must print ALL WRITES SEMANTICALLY CORRECT). \\
    Full campaign & Primary matrix ($\approx$~hours): \texttt{INDEP=1 REPEATS=25
      DURATION=12 WARMUP=15 node bench/runner.mjs} $\rightarrow$ \texttt{results/raw.json};
      all secondary experiments add up to $\approx$25~h wall-clock. \\
    Regenerate from raw & $\approx$minutes, \emph{no database}: the standalone no-database renderers rebuild their mapped tables and figures from \texttt{results/*.json}, then
      \texttt{npm run sync:tables} and \texttt{make}; \texttt{MANIFEST.md} explicitly marks
      authored tables, the one pre-generated statement-log table, and seven run-coupled renderers. \\
    Verification status & Release v1.12.13.s checksum manifest verifies 37 result JSON files. In a fresh copy without dependencies or build outputs, the complete standalone-renderer chain rebuilt both table directories byte-for-byte; this was an author-run offline reconstruction, not a benchmark rerun. The historical immutable-v1.12.9 log verifies 35/35 inputs and 45/50 then-committed tables. The revised measurements have not been independently reproduced and still require the reference engines. \\
    Entry point & \texttt{REPRODUCE.md} (smoke test, full matrix, clean-room reproduction
      from the tarball); the per-table data-and-generator map is in \texttt{MANIFEST.md}. \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[htbp]
  \centering
  \footnotesize
  \caption{Scope of every experiment: the access pattern(s), engine(s), number of
    access-layer implementations, and repeated runs each covers. The \emph{primary
    comparison} (throughput and closed-loop p99) is the full five-pattern, two-engine
    matrix; every other experiment is a limited condition, most restricted to the
    deep fetch, and each conclusion is scoped accordingly. ``Layers'' counts
    implementations (the nine access layers plus, where present, the two tuned native
    baselines); a representative subset is used where a caption says so. Engine
    entries name where the experiment was run (\texttt{pg} is PostgreSQL-only,
    \texttt{mysql2} MySQL-only).}
  \label{tab:scope}
  \begin{tabular}{@{}p{0.34\textwidth} p{0.13\textwidth} p{0.11\textwidth} c p{0.15\textwidth}@{}}
    \toprule
    Experiment & Pattern(s) & Engines & Layers & Runs \\
    \midrule
    Primary matrix: throughput \& closed-loop p99 (main paper) & all five & PG, MySQL & 11 & 25 independent \\
    Same-SQL raw-path contrast (Table~\ref{tab:sameplan}) & deep fetch & PG, MySQL & 9 & median of 10 \\
    Capacity-denominator sensitivity (Table~\ref{tab:capacity_sensitivity}) & deep fetch & PG, MySQL & 8 & archived-data re-expression \\
    Open-loop CO-corrected tail (Tables~\ref{tab:openloop},~\ref{tab:openloop_mysql}) & deep fetch & PG, MySQL & 5 & rate sweep \\
    Utilization-controlled tail (Tables~\ref{tab:utilization},~\ref{tab:utilization_mysql}) & deep fetch & PG, MySQL & 8 & median of 5 \\
    Equal-CPU, 1/2/4 cores (Table~\ref{tab:equalcpu}), exploratory & deep fetch & PG & 3 & median of 3 \\
    Node-cluster multi-worker (Table~\ref{tab:cluster}) & deep fetch & PG, MySQL & 3 & median of 3 \\
    Pool-size sweep, 1--50 (Tables~\ref{tab:poolsize},~\ref{tab:poolsize_mysql}) & deep fetch & PG, MySQL & 4 & median of 3 \\
    Mixed read/write, 90:10 \& 70:30 (Table~\ref{tab:mixed}) & deep fetch, insert & PG, MySQL & 5 & median of 5 \\
    Relaxed vs default durability (Table~\ref{tab:durability}) & insert & PG, MySQL & 11 & 25 def.\ $+$ 10 relaxed \\
    Wait-event flush decomposition (Table~\ref{tab:waitevents}) & insert & MySQL & 3 & 4000 inserts/layer \\
    Transactional multi-statement write (Table~\ref{tab:txn_write}) & \texttt{POST /threads} & PG & 5 & median of 5 \\
    Longer-window (60\,s) tail (Table~\ref{tab:taillong}) & deep fetch & PG, MySQL & 11 & 10 runs \\
    Canonical-constructor microbenchmark (Table~\ref{tab:canonicalization_cost}) & all read shapes & none & shared code & 30 blocks \\
    Post-restart cold cache (Table~\ref{tab:postreboot}) & deep fetch & PG, MySQL & 3 & vs primary median \\
    Alternative eager-loading (Table~\ref{tab:altloading}) & deep fetch & PG, MySQL & 3 & median of 5 \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Construct validity of the documentation-selected treatment}
\label{sup:construct}
The \emph{documentation-selected} rule is an intentionally artificial, predeclared treatment-selection policy, not a model of what a typical developer would choose. It selects the first relations API on the version-compatible official page under the stated conflict rule. For auditability, the artifact preserves the exact response returned by the nearest available Wayback capture at or before the freeze for each official URL; the committed HTML, capture timestamp, and SHA-256 are indexed by \texttt{experiments/documentation-snapshots/manifest.json}, and the per-layer decision appears in \texttt{METHODOLOGY.md} and Table~\ref{tab:adapter_choices}. A capture establishes page state only at its timestamp, not continuous lack of change through the freeze. This provides reproducibility and auditability, not evidence of production frequency or optimality. Table~\ref{tab:construct} records this per layer. Drizzle is the one borderline case: it exposes both a SQL-style query builder (the selected join) and a higher-level relational-query API that its documentation also presents, so we selected the builder join --- comparable with the query-builder tier --- and record the relational-query API as the documented alternative.
Where the documented alternative is a byte-identical drop-in, the loading-strategy sensitivity
check (Table~\ref{tab:altloading}) adds an empirical signal: for the join-default ORMs (Sequelize,
MikroORM) the documentation-selected default is the \emph{faster} of the two strategies, so a
layer's deep-fetch deficit does not appear to be an artifact of a poor default; for Objection the documented
default (batched select-in) is the slower option on MySQL, which we disclose rather than correct; the
Drizzle, Prisma, and TypeORM alternatives were not evaluated (TypeORM's errors on the ordered
deep fetch), itself a portability caveat. The construct is therefore a reproducible policy-defined treatment, not a practitioner persona and not a claim about performance-tuned expert usage or measured production frequency; RQ1 is scoped accordingly (main text, Threats to Validity).

\begin{table}[htbp]
  \centering
  \footnotesize
  \caption{Construct-validity corroboration for the documentation-selected deep-fetch treatment:
    the selected eager-loading API, whether it is the library's canonical (documentation-primary)
    relations mechanism, and the empirical loading-strategy signal where the documented
    alternative is a byte-identical drop-in (Table~\ref{tab:altloading}).}
  \label{tab:construct}
  \begin{tabular}{@{}l l p{0.30\textwidth} p{0.30\textwidth}@{}}
    \toprule
    Layer & Selected API & Canonical relations mechanism? & Loading-strategy signal \\
    \midrule
    \texttt{pg}, \texttt{mysql2} (+tuned) & hand-written JOIN & Yes --- the reference way to fetch a graph with a driver; no higher-level relations API applies & n/a (hand-written SQL) \\
    \texttt{knex} & builder \texttt{.join()} & Yes --- the query builder's documented join & n/a (hand-written SQL) \\
    \texttt{drizzle} & builder join & Documented; the relational-query API is the alternative & alternative (relational-query API) not evaluated \\
    \texttt{prisma} & \texttt{include} & Yes --- Prisma's primary relation-query API & alternative (join strategy) not evaluated \\
    \texttt{sequelize} & \texttt{include} (join) & Yes --- Sequelize's canonical eager loading & default \emph{faster} than select-in \\
    \texttt{typeorm} & \texttt{relations} & Yes --- TypeORM's documented relations option & alternative errors on the ordered deep fetch (not evaluated) \\
    \texttt{objection} & \texttt{withGraphFetched} & Yes --- Objection's flagship eager-loading & default (select-in) slower than join on MySQL (disclosed) \\
    \texttt{mikroorm} & \texttt{populate} (join) & Yes --- MikroORM's canonical eager loading & default \emph{faster} than select-in \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Results roadmap}
\label{sup:outcomes}
Table~\ref{tab:outcomes} classifies every result in the paper by its analysis role --- which
outcomes carry the confirmatory, paired inference (\emph{primary}), and which are reported
descriptively as controls or sensitivity analyses (\emph{secondary} and \emph{exploratory}) that
constrain or contextualize the primary results rather than adding confirmatory hypotheses. It is the
navigational companion to the reproduction map in \texttt{experiments/MANIFEST.md}, which ties each
table to its generator and input data. The two co-primary deep-fetch regimes it lists under RQ1 ---
documentation-primary and performance-conscious --- are now measured and reported in the main
text's Results section, no longer relegated to the supplement.

\input{\tabledir outcomes}

\section{Per-pattern capacity and utilization at the operating point}
\label{sup:scaling-patterns}
The operating-point protocol requires that capacity be identified so the fixed operating point can be
read against it (main text, Study Design). The deep-fetch concurrency sweep
(Section~\ref{sup:scaling}, Figure~\ref{fig:scaling}) does this in per-layer detail for one pattern;
review point 6.5 asks whether the other four patterns, reported only at the fixed 50-connection
demand, sit at comparable utilization. To answer it we swept every access layer over a connection
ladder (1, 4, 8, 16, 32, 50, 100, 200) on \emph{all five} patterns and both engines, medianed over
two runs per point after a warm-up, and for each layer computed its saturating throughput (the ladder
maximum), the utilization at 50 connections (throughput at 50 connections divided by that maximum),
and the knee (the smallest connection count reaching 95\% of the maximum).

Table~\ref{tab:scaling_patterns} reports the result. The knee lies at or below 50 connections for
every pattern on both engines, so the fixed 50-connection operating point places all five patterns at
high utilization of their own capacity: cross-pattern comparisons of p99 and spread at that point are
therefore made at comparable (high) relative utilization, not at unknown or widely different
fractions of capacity. The binding constraint is the ten-connection pool shared by every pattern
(main text, Controls), which is why the knee clusters well below 50 connections regardless of the
pattern's absolute throughput. This extends the capacity-identification stage of the protocol from
the deep fetch to all five patterns; the equal-utilization open-loop tail
(Section~\ref{sup:utilization}) and the exploratory equal-compute check (Section~\ref{sup:equalcpu}),
which each need a per-layer saturating-throughput \emph{target}, remain demonstrated on the deep
fetch, the pattern on which the layer spread is largest.

\input{\tabledir scaling_patterns}

\section{Paired significance of the deep-fetch ranking}
\label{sup:significance-deep}
The main text foregrounds paired \emph{effect sizes} for the deep-fetch ranking --- geometric-mean
per-replicate throughput ratios and the fraction of replicates in which the faster layer wins ---
because they, not the significance tests, carry the ordering. Table~\ref{tab:significance} reports the
post-hoc adjacent-pair tests as a descriptive companion: since the ranking itself selects which pairs
are compared, the inference is conditional, so the permutation and Wilcoxon results are not treated as
confirmatory. The bootstrap intervals are within-campaign (main text, Study Design), not general over
deployment environments.

\input{\tabledir significance_deep_fetch}

\section{Application-tier CPU trade-off}
\label{sup:cpu-tradeoff}
Figure~\ref{fig:cpu_tradeoff} plots each layer's application-tier CPU against its throughput on the
deep fetch (the trade-off discussed in the main text's Resource-footprint paragraph): every layer
draws about one application core, and the native driver leads on throughput and on requests per CPU-second (application-tier and
combined), consistent with its leaner SQL; these CPU shares, measured at unequal throughput, do not
identify where per-request work occurs.

\input{\tabledir fig_cpu_tradeoff}

\section{Retrospective application of the protocol to prior benchmarks}
\label{sup:protocol-retro}
Table~\ref{tab:protocol_retro} applies the protocol retrospectively to the eight prior benchmarks
located in the sources searched. The exercise is purely analytical: we re-ran nothing and dispute
none of their reported numbers. For each study we name the protocol stages that are absent and state,
in one line, why the study's headline conclusion is not interpretable as a general access-layer claim
without them. The point is not that these studies erred, but that naming the missing stages localizes
exactly what each study's measurements can and cannot support --- and, for the one engine-only study,
why the access-layer unit of comparison is absent by design.

\input{\tabledir protocol_retro}

\section{Randomized differential semantic-equivalence check}
\label{sup:semantic-equivalence}
The fixed 12-probe gate (\texttt{bench/verify.mjs}) establishes finite pre-measurement conformance;
the main text (Study Design) also reports a broader, property-based level. Because the benchmark
dataset is deterministic and finite, testing far more inputs is inexpensive relative to the
measurement campaign. A second gate (\texttt{bench/verify-property.mjs}) generates a fixed-seed
sample of one thousand random post IDs and one thousand random author IDs across the full key range,
together with an explicit edge set (nonexistent, boundary, negative, and $2^{31}{-}1$ IDs; empty and
boundary keyset pages), and asserts that every adapter is byte-identical to the native-driver oracle
on all five read methods. Table~\ref{tab:semantic_equivalence} summarizes the coverage: 3{,}800
distinct inputs per adapter, 30{,}400 adapter-versus-baseline comparisons per engine, 60{,}800 across
both engines, with zero divergences. This does not prove equivalence on every possible input --- that
remains infeasible for arbitrary code --- but it broadens the finite check by more than two orders of
magnitude over the 12-probe gate and exercises the boundary and error-path inputs most likely to
reveal an input-dependent divergence. The sample size is a script parameter, so broadening the randomized
check further is a re-run at larger $N$ against the deterministic dataset, not a redesign.

\input{\tabledir semantic_equivalence}

\section{The five access patterns}
\label{sup:patterns}
Table~\ref{tab:patterns} lists the five HTTP endpoints of the benchmark workload, the access
pattern each realizes, and what each is intended to stress. The main text (Study Design) describes
the patterns in prose; this table is the compact reference.

\input{\tabledir patterns}

\section{Protocol compliance levels and their coverage}
\label{sup:protocol-compliance}
The main text (Study Design) names five protocol compliance levels; Table~\ref{tab:protocol_compliance}
records precisely which workload cells satisfy each in this case study. The mandatory validity core
and capacity characterization hold across all five patterns on both engines; the three
extensions --- standardized-strategy, utilization-controlled, and resource-normalized --- are
demonstrated on the deep fetch, the most consequential pattern, as a representative-point
instantiation rather than a claim of full-matrix coverage. Stating the levels and their coverage
explicitly is what makes the protocol reusable: a later study can report its own compliance level per
cell against the same ladder.

\input{\tabledir protocol_compliance}

\section{Predefined contrasts against the native-driver reference}
\label{sup:native-contrasts}
Selecting adjacent pairs \emph{after} seeing the ranking makes those tests conditional on the observed
ordering, so their $p$-values add little (main text, Measurement stability; the adjacent-pair table is
Table~\ref{tab:significance}). A predeclared contrast set avoids this. Table~\ref{tab:native_contrasts}
fixes the reference (the native driver) and the contrasts (each portable layer against it) in advance,
independently of the ranking, and reports for each layer on the deep fetch the geometric-mean paired
ratio of native-to-layer throughput, a within-campaign 95\% paired-bootstrap interval, and the paired
dominance. Every portable layer trails the native driver in all 25 replicates on both engines
(dominance $1.00$), and the ratios range from $1.48\times$ (\texttt{knex}, PostgreSQL) to $7.16\times$
(\texttt{mikroorm}, PostgreSQL) --- the same spread the main text reports, now as predeclared contrasts
rather than post-hoc pairwise tests. The contrasts are within-campaign, not deployment-general.

\input{\tabledir native_contrasts}

\section{Run-to-run p99 spread}
\label{sup:p99-spread}
Each cell's reported p99 is the median of 25 run-level p99 values, and each run-level p99 is itself
estimated from roughly the top 1\% of that run's requests, so a single run's p99 is noisy.
Figure~\ref{fig:p99_spread} plots all 25 per-run p99 values behind each deep-fetch cell (PostgreSQL)
with the median marked, so the run-to-run stability of the estimate is visible rather than summarized
by a single number: the spread is tight for the fast layers and wider for the slow ones, consistent
with the coarser p99 estimate at higher latency, and the medians preserve the reported ordering. The
load generator reports latency percentiles rather than raw per-request latencies, so a full
HDR-histogram reconstruction is left to future work; the median-of-run-level-p99 estimator, its
within-campaign bootstrap interval, and the longer-window (60~s) sensitivity check
(Section~\ref{sup:taillong}) are the mitigations used here.

\input{\tabledir fig_p99_spread}

\section{Components moved by the same-SQL standardized contrast}
\label{sup:samesql-components}
The same-SQL (raw-path) standardized contrast (main text, RQ1) is a \emph{compound} intervention: replacing each
layer's documentation-selected relational-loading path with the same hand-written SQL through its raw
facility changes several mechanisms at once. Table~\ref{tab:samesql_components} enumerates them. The
residual spread it leaves is a standardized-SQL contrast --- not a bound on the combined effect, and
not attributable to any single component, in particular not to SQL formulation or query count.

\input{\tabledir samesql_components}

\section{Deep-fetch tail: equal closed-loop demand and matched utilization}
\label{sup:tailregimes}
The high-load p99 under equal closed-loop demand (main text, Tail latency) and the p99 at matched
fractions of each layer's own capacity read the deep-fetch tail two ways. Table~\ref{tab:tail_regimes}
shows both per layer, so the large high-load gap (\texttt{pg}~20 to \texttt{mikroorm}~116~ms on
PostgreSQL) is seen to be a capacity-and-queueing effect that largely dissolves to a $2$--$5$~ms band
at matched utilization, rather than a sub-saturation latency difference.

\input{\tabledir tail_regimes}

\section{Standalone cost of canonical response construction}
\label{sup:canonicalization}
The shared constructors enforce comparable output, but could themselves mask layer cost if they performed substantial graph work. They only copy selected fields, normalize numbers, booleans, and ISO timestamps, and map arrays that the adapter has already grouped; they perform no query, join, regrouping, sorting, or cache access. Table~\ref{tab:canonicalization_cost} reports their isolated cost. Even the largest fixture is 0.146\% of the smallest corresponding primary-campaign HTTP p50; the ten-comment deep-fetch forms are 0.078\%. These are single-process development-host measurements and a scale comparison, not a decomposition of request time.

\input{\tabledir canonicalization_cost}

\section{Sensitivity of the matched-utilization capacity denominator}
\label{sup:capacity-sensitivity}
The utilization experiment used the 25-run median throughput at 50 connections as a
predefined capacity proxy. Table~\ref{tab:capacity_sensitivity} re-expresses the same offered
loads against a separate empirical denominator: the maximum observed in the full concurrency
sweep. The proxy is 93.5--102.3\% of that maximum across the sixteen layer--engine cells, so the
nominal 50\% and 70\% loads become 46.8--51.2\% and 65.5--71.6\%. This sensitivity analysis
changes no p99 observation and keeps both interpretation bands below saturation.

\input{\tabledir capacity_sensitivity}

\end{document}
